---
title             : "Application of Close-kin mark–recapture (CKMR) to Protogynous Fishes"
shorttitle        : "\\mbox{}"

author: 
  - name          : "Claudia Friess"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : ""
    email         : "elasmophile@gmail.com"
    role:         
  - name          : "Eric C. Anderson"
    affiliation   : "2,3"
    role:
  - name          : "Lisa E. Ailloud"
    affiliation   : "4"
    role:
  - name          : "David S. Portnoy"
    affiliation   : "5"
    role:
  - name          : "Susan K. Lowerre-Barbieri"
    affiliation   : "1,6"
    role:

affiliation:
  - id            : "1"
    institution   : "Florida Fish and Wildlife Research Institute, St. Petersburg, FL, USA"
  - id            : "2"
    institution   : "NOAA, National Marine Fisheries Service, Southwest Fisheries Science Center, Fisheries Ecology Division, Santa Cruz, CA, USA"
  - id            : "3"
    institution   : Department of Biology, Colorado State University, Fort Collins, CO, USA
  - id            : "4"
    institution   : "NOAA, National Marine Fisheries Service, Southeast Fisheries Science Center, Charleston, SC, USA"
  - id            : "5"
    institution   : "Marine Genomics Laboratory, Texas A&M University, Corpus Christi, Corpus Christi, TX, USA"
  - id            : "6"
    institution   : "University of Florida, Fisheries and Aquatic Sciences, Gainesville, FL, USA"

authornote: ""


abstract: |
  Abundance estimates for marine fish are difficult to obtain and typically rely on 
  stock assessment models that assume only females matter to productivity. Many fished 
  species, however, are protogynous hermaphrodites for which preferential harvest of certain size classes can drastically 
  skew sex ratios. This raises concerns about sustainability, especially if managed based 
  on female-only reference points. There is a growing interest in integrating Close-Kin Mark-Recapture 
  (CKMR) into stock assessments to reduce uncertainty in abundance estimates, but this 
  has not previously been done for protogynous hermaphrodites. Here we derive the CKMR kinship probabilities 
  for two relationship types---Parent-Offspring Pairs (POP) and Half-Sibling Pairs (HSP)---
  and include them in an integrated stock assessment framework
  for estimating abundance. Using simulation, we evaluate the robustness of this approach to 
  uncertainty in male contribution to reproductive success and kin pair parental sex assignment misspecification. 
  We model our simulation after Gag, *Mycteroperca microlepis*, whose population in the Gulf of Mexico
  has experienced large 
  declines and considerable shifts in sex ratios over the last half century.
  We show that CKMR can produce unbiased abundance and sex ratio estimates only 
  if the relative contribution of male and female age classes to reproductive success 
  is not misspecified and sex of the parent in POPs or shared parent in HSPs is identified
  correctly. We also highlight the potential for CKMR data to estimate 
  parameters of the sex transition and male reproductive contribution functions in an 
  integrated modeling framework.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "simulation, mtDNA, reproductive resilience, integrated stock assessment, hermaphrodites, genetic tagging"

bibliography      : ["citations_abbrev.bib", "r-references.bib"]
csl               : "elsevier-harvard.csl"

floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no
always_allow_html : true

classoption       : "man"
output            : papaja::apa6_pdf
header-includes   : 
  - \usepackage{color}
  - \usepackage{tipa}
  - \usepackage{wasysym}
  - \usepackage{fancyhdr}
  - \usepackage{mathrsfs}
  - \thispagestyle{empty}
  - \usepackage{hanging}
  - \captionsetup[figure]{labelfont=bf}
  - \captionsetup[table]{labelfont=bf, labelsep = period, textformat = simple}
  - \raggedbottom
  - \newcommand{\tprob}{P^{F\rightarrow M}}
  - \newcommand{\wt}[1]{\omega_#1}  % recruitment deviations
  - \newcommand{\rbar}{\bar{R}}  % mean recruitment
  - \newcommand{\fbarr}{{\bar{F}_\mathrm{rec}}}  % mean fishing mortality from rec fishery
  - \newcommand{\frecdev}[1]{fdev^\mathrm{rec}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\fbarc}{{\bar{F}_\mathrm{com}}}  % mean fishing mortality from commercial fishery
  - \newcommand{\fcomdev}[1]{fdev^\mathrm{com}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\Mage}[1]{\dot{M}_#1}  % age-specific instantaneous natural mortality
  - \newcommand{\vulr}[1]{\dot{v}^\mathrm{rec}_#1}  % age-specific vulnerability to recreational fishery
  - \newcommand{\vulc}[1]{\dot{v}^\mathrm{com}_#1}  % age-specific vulnerability to commercial fishery
  - \newcommand{\dFec}[1]{\dot{Fec}_#1}    % age-specific relative fecundity
  - \newcommand{\Af}[1]{\dot{Mat}_#1}    % age-specific fraction of females mature
  - \newcommand{\Pfm}[1]{\dot{P}^{\mathrm{F}\rightarrow\mathrm{M}}_#1}    % age-specific prob of transitioning from female to male
  - \newcommand{\Imp}[1]{\dot{P}^\mathrm{M}_#1}    % equilibrium proportion of males amongst age a fish
  - \newcommand{\It}[1]{\mathring{I}_{#1}}   % recruitment index in year t
  - \newcommand{\pacom}[2]{\mathring{p}^\mathrm{com}_{#1,#2}}   % relative abundance of age a females in year t
  - \newcommand{\pacr}[3]{\mathring{p}^\mathrm{ckmr}_{#1,#2,#3}}   % relative abundance of age a males in year t
  - \newcommand{\Z}[2]{Z_{#1,#2}}   % total instantaneous mortality at time t for age a fish 
  - \newcommand{\AS}[2]{S_{#1,#2}}   % total annual survival
  - \newcommand{\lxf}[1]{s_{#1}}     % survivorship
  - \newcommand{\Nf}[2]{N^\mathrm{fem}_{{#1}, {#2}}}  % number of females of age a at time t
  - \newcommand{\Nm}[2]{N^\mathrm{mal}_{{#1}, {#2}}}  % number of males of age a at time t
  - \newcommand{\K}[4]{\mathring{K}^{\mathrm{#1},{{#4}}}_{{#2},{#3}}}
  - \newcommand{\KHS}[3]{\mathring{K}^{\mathrm{HSP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSs}[2]{\mathring{K}^{\mathrm{HSP},{{#2}}}_{{#1}}}
  - \newcommand{\KPOPT}[3]{\mathring{K}^{\mathrm{POP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSD}[2]{\mathring{K}^{\mathrm{HSP}^{\neq _{a,b}}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with different mtDNA 
  - \newcommand{\KHSS}[2]{\mathring{K}^{\mathrm{HSP}^{=_a}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOP}[3]{\mathring{K}^{\mathrm{POP}^{#3}}_{{#1},{#2}}}  
  - \newcommand{\KPOPF}[2]{\mathring{K}^\mathrm{POP_F}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOPM}[2]{\mathring{K}^\mathrm{POP_M}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA 
  - \newcommand{\HSPD}{\mathrm{HSP}^{\neq}}
  - \newcommand{\HSPS}{\mathrm{HSP}^=}
  - \newcommand{\POPD}{\mathrm{POP}^{\neq}}
  - \newcommand{\POPS}{\mathrm{POP}^=}
  - \newcommand{\KNot}[2]{\mathring{K}^\mathrm{Not}_{{#1},{#2}}}     % number of pairs that are neither HSD or HSS
  - \newcommand{\KNote}[2]{\mathring{K}^{\mathrm{Not}=_a}_{{#1},{#2}}}
  - \newcommand{\KNotu}[2]{\mathring{K}^{\mathrm{Not}{\neq}_{a,b}}_{{#1},{#2}}}   
  - \newcommand{\Hmt}{H_\mathrm{mt}}  % mtDNA haplotype diversity
  - \newcommand{\Hmts}{H_\mathrm{mt}^{\mathrm{OM}}}  % simulated mtDNA haplotype diversity
  - \newcommand{\Hmte}{H_\mathrm{mt}^{\mathrm{EM}}}  % mtDNA haplotype diversity assumed in the EM
  - \newcommand{\Mat}{\mathrm{Mat}}
  - \newcommand{\Fec}{\mathrm{Fec}}
  - \DeclareMathAlphabet{\mathbbold}{U}{bbold}{m}{n}
  - \newcommand{\comment}[1]{\textcolor{red}{#1}}
  - \newcommand{\ERRO}{\mathrm{ERRO}}
  - \newcommand{\ETRO}{\mathrm{ETRO}}
  - \newcommand{\sH}{\mathscr{H}}
---


```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
library(gridExtra)
library(dplyr)
library(here)

# caching for quick knitting.  
#    set cache = FALSE to not cache
#    set cache.extra to a new value to remove all caches and re-evaluate
knitr::opts_chunk$set(cache.extra = 3, cache = FALSE) 


```

```{r tables_figs, include = FALSE}
source(here("code", "analysis", "results_tables.R"), local = knitr::knit_global())
source(here("code", "analysis", "results_figures.R"), local = knitr::knit_global())
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

# 1. Introduction

Generating accurate estimates of marine fish population abundance is one of the most challenging 
yet essential mandates of fisheries science. To meet it, ever-more sophisticated stock assessment
models, capable of integrating a variety of data sources, are being developed [@punt2020essential].
At the core, however, these models are based on the same basic fish population dynamics 
equations developed for gonochoristic species decades ago [@beverton1957dynamics; @hilborn1992quantitative]. Similarly unchanging
has been the central overarching goal of fisheries management: optimize resource exploitation by balancing current 
extraction against future productivity, or reproductive potential. 
The key, then, is how we define, both conceptually and operationally, "reproductive potential". In terrestrial mammals, it is driven by the 
number and fecundity of the females. This paradigm has been adopted for marine fish in the form of measures
like spawning stock biomass (SSB) or total egg production, although problems with this approach are increasingly 
recognized [@sharma2019recruitment]. The reproductive resilience paradigm posits that marine fish reproductive strategies 
are unlike those seen on land and evolved to achieve lifetime reproductive success and maintain population 
stability within highly variable environments [@tringali2023reproductive; @lowerre2017reproductive]. 
These reproductive strategies include extreme fecundity, external fertilization of pelagic eggs, 
offspring survivorship affected by where and when fish spawn, and sequential hermaphroditism.
Both protandry (reproducing as males first and then as females) and protogyny (reproducing as females first and changing sex to male) 
occur in marine fishes, but protogyny is most common [@pla2022switches], especially in warmwater
systems [@lowerre2023unified].

Delayed male maturation, such as seen in protogynous species, is predicted to evolve when there
is male competition for females. Sex allocation theory predicts that adult sex ratios are driven
by processes that result in optimal mean reproductive success for the population [@charnov1982theory; @warner1988sex] and the size 
advantage model predicts that a fish will change sex when reproductive success at age is greater for
the terminal sex [@ghiselin1969evolution; @warner1975reproductive; @munoz2003new]. 
Fisheries target larger animals, truncating population age and size structure [@barnett2017old], 
which disproportionately increases the risk of overexploitation for males in protogynous species. Protogynous species 
vary in their ability to adapt to heavy fishing pressure and the selective removal of males. Those that form harems 
with an alpha male typically exhibit rapid sex change which can mitigate fishing-induced male depletion [@warner1982mating]. By contrast, species that form leks or spawning aggregations may be more vulnerable because their sex-change dynamics and response 
to male removal remain poorly understood, and evidence from exploited stocks suggests frequent male 
depletion under fishing pressure [@lowerre2020testing; @alonzo2004effects; @alonzo2005sex; @provost2015impacts].
Many protogynous species support important fisheries, 
such as the keystone predator California sheephead, *Bodianus pulcher*, [@alonzo2004status; @hamilton2015exploitation] and 
heavily-fished serranid groupers in the Gulf of Mexico (GOM) and Caribbean Sea.

Two current areas of active research in fisheries stock assessment are defining and measuring 
reproductive potential for protogynous hermaphrodites [@alonzo2008importance; @brooks2008stock;
@alonzo2005sex] and incorporating genetic data into integrated models [@punt2020essential; @punt2024including; @fisch2025expected].
Of particular interest is Close-Kin Mark-Recapture (CKMR; @bravington2016close, 
@julius2001allele; @nielsen2001statistical), a method that
uses genetic data to identify kin pairs in a sample and treats those
kin pairs as "recaptures" in a mark-recapture-like framework [see @casas2023review for a 
review of CKMR and its application to fisheries].
CKMR has been used to estimate absolute abundance for a number of marine and freshwater fish 
species with a range of reproductive strategies, including the
gonochoristic southern Bluefin tuna, *Thunnus thynnus*, [@bravington2016absolute; @davies2020next], 
semelparous Chinook salmon, *Oncorhynchus tshawytscha*, [@rawding2014genetic], and elasmobranchs which have a fixed-ratio 
reproductive strategy [@hillary2018genetic; @delaval2023evaluating].
Integrating CKMR into a stock assessment requires including the CKMR pseudolikelihood
[@bravington2016close] as one component in the stock assessment model.
This CKMR pseudolikelihood relates the observed number of genetically identified closely related pairs
(e.g., parent-offspring or half-sibling-pairs) in a sample to the
number expected given the size of the population. The expected
number of kin pairs is found by deriving a set of _kinship probabilities_, each one being
the probability that a pair of samples are close kin, given the size of the population and
additional information that describes the likelihood of an individual surviving and
reproducing over time relative to other individuals with similar demographic characteristics. 

To date, nearly all CKMR applications have been to gonochoristic species in which 
it has been either assumed or observed that fecundity drives reproductive success 
and that this success scales equally between males and females. The life history of protogynous hermaphrodites 
complicates application of CKMR because 
sex ratios are typically skewed toward females, and, amongst half-sibling
pairs, one fish's mother could be another's father. In fact, CKMR kinship probabilities 
for protogynous hermaphrodites have not previously been derived. Additionally, it is 
unknown whether traditionally-used heuristics for the CKMR sample sizes required
to produce accurate and precise abundance estimates in gonochoristic species are 
also appropriate for protogynous hermaphrodites. Finally, even though current best
practice in stock assessments of protogynous hermaphrodites is to assume male and
female SSB contribute equally to reproductive potential [@brooks2008stock], it has not been determined 
whether this is, in fact, the case, nor do we know if and how a declining male sex ratio
affects population productivity.

In this study, we derive the kinship probabilities for protogynous hermaphrodites,
extending the work of @bravington2016close to use the frequency of kinship types 
observed in genetic data to estimate abundance and demographics in protogynous 
species. We then conduct a simulation study, modeled after Gulf of Mexico Gag, *Mycteroperca microlepis*, 
to investigate the robustness of these estimators to uncertainty in 
key reproductive and genetic parameters. 
One focal question was whether the CKMR data contain sufficient information to estimate 
the shapes of the sex-transition and male age-specific reproductive-output functions used 
in the CKMR likelihood. Gag are of particular interest because their spatial ecology, 
sexual segregation, and lek-like reproductive unit have resulted in extreme male scarcity 
under heavy fishing pressure and limited plasticity in sex change [@lowerre2020testing]. This has raised concerns 
about depensation at low abundance due to sperm limitation. Our study underscores 
the importance of using simulation to evaluate the effectiveness of a CKMR study prior to 
implementation, and identify research priorities to resolve uncertainties most likely to 
obscure or invalidate results. While this study focuses on Gag, the CKMR estimators and general 
simulation framework are adaptable to a wide range of protogynous fishes.

# 2. Methods

## 2.1. Overview

The study can be broken down into the following core steps: 

1. Operating Model (OM): We use an individual-based operating model (Section 2.4) 
to generate pedigrees for a population of protogynous hermaphrodites. Each simulation 
creates millions of fish in silico with known birth and death dates, sex, mortality 
causes, and parentage. The simulated population experiences increasing fishing mortality, 
resulting in population decline and growing sex-ratio skew, consistent with trends 
observed in the real-world Gag population. 
2. Sampling Model: We sample individuals from the OM (Section 2.5) and generate both traditional fisheries 
data (i.e., a relative abundance index and composition data) and CKMR kin pairs 
(parent-offspring pairs (POPs) and half-sibling pairs (HSPs)) to serve as inputs 
to the estimation model (Section 2.6). 
3. Estimation Model (EM): We use a simple, integrated fisheries stock-assessment model 
to estimate sex specific abundance. By comparing these estimates against the ‘true’ 
abundance generated by the OM, we evaluate model performance across two scenario 
groups (Section 2.7): one focused on the effect of parental sex assignment misspecification, 
and another exploring uncertainty surrounding male contribution to reproductive success.
4. Model Evaluation: We quantify and explore the accuracy and variability in estimated 
abundance and male sex ratio under different scenarios (Section 2.8). 


## 2.2. Model Species

Our operating  model was based on the life history of Gag in the GOM. 
Gag live to be over 30 years old and 
mature around age 4. Gag exhibit clear depth preferences with life stage, sex, 
and spawning. They use estuarine nursery habitat as young of the year, 
then move to nearshore sites with structure as young females (~ages 1 to 4) 
and offshore to spawn as they mature. Individuals undergoing sex transition occur before, 
during, and after the spawning season and in both nearshore, all-female 
pre-spawning aggregations and in offshore spawning grounds [@lowerre2020testing].
Gag spawn at the shelf edge from February to mid-April in lek-like reproductive units 
and produce pelagic eggs. Larvae settle 
in estuaries after 35 to 45 days [@fitzhugh2005spatial; @lowerre2020testing].
Males are thought to remain in deeper waters on the shelf edge year-round whereas some females 
leave the spawning grounds, presumably returning to shallower waters. 
All Gag start life as females, and
a subset of them change sex to male as they age. Although a male as young as four years 
old has been observed, age at 50% male is approximately 12 years. 

Gag are highly sought after by both commercial and recreational fisheries in 
the GOM. The majority of the quota is allocated to the recreational 
sector, and estimates of recreational landings are highly uncertain [@SEDAR2022; @NOAA2023]. 
Gag have been estimated to be overfished and experiencing overfishing almost 
consistently since the late 1960’s [@SEDAR2022]. The spawning stock biomass estimated 
in the 2022 stock assessment was about 2% of unfished levels, and the sex ratio has declined from an 
estimated 32% male in unfished conditions to between 1 and 5% male in recent years 
[@lowerre2020testing; @SEDAR2022], with little observed change in the size and age at sex transition [@lowerre2020testing]. 
Additionally, skipped spawning is relatively common in female Gag,
and is thought to be linked to low male abundance [@lowerre2022spawning]. 
It is unknown to what extent sperm limitation due to reduced male availability 
is responsible for the continued decline of the Gag population in the GOM [@lowerre2021gag]. 

## 2.3. CKMR Kinship Probabilities

We limit this study to HSPs and POPs and
assume that these can be identified without error. We furthermore assume that ages of 
sampled individuals are known without error, sampling is conducted on dead
individuals, and age-specific survival is independent of sex.  
The full CKMR model derivation can be found in Appendix A. The probabilities of observing
HSPs (Equations A1-A7) account for the unknown age and sex of the shared parent as
well as the possibility that the first-born's mother could be the second-born's father.
Similarly, the probabilities of observing POPs (Equations A9-A13) account for the fact that
the parent, if it was male at the time of sampling, could have been female when it
gave birth to the offspring in the pair. Sex change is included via the age-specific
probability of a female changing sex to become male, $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
(see Table \@ref(tab:table1) for this and other notation used in this paper). 

The ability to identify---or at least infer---the sex of a sampled individual's parent is crucial for 
applying the model, and in our model this is accomplished by sequencing the fish's
mitochondrial DNA (mtDNA) which is exclusively inherited from the female parent. For the purposes
of the simulation, we assumed that different mtDNA haplotypes were equifrequent
(see Supplementary Material for more discussion on the topic),
so that we were able to parameterize the mtDNA haplotype frequencies simply by $H_\mathrm{mt}$---the
probability that two randomly sampled mtDNAs from the population
are different in sequence. This quantity is also commonly called the mtDNA
haplotype diversity [@avise2012molecular], a quantity that can be estimated by 
typing the mtDNA of a large sample of individuals.
Parameterizing the simulations in terms of it allowed
us to avoid simulating mtDNA sequences
in all simulated individuals and also simplified the CKMR likelihood used to analyze the simulations 
(see Supplementary Material). An $\Hmt$ of 1 would indicate a 100% certainty
that two HSPs with the same mtDNA sequence shared a mother, while any value <1 means there is
a chance (which equals $1 - \Hmt$) that their mtDNA sequence is the same even though they did not
share a mother (i.e., they inherited their mtDNA sequence from a common ancestor other than their mother). 
We incorporate uncertainty due to $\Hmt$ into our CKMR model
equations (Equations A8 and A14) and in the simulation, we explore the effects of 
misspecifying both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\Hmt$ in the estimation model (Section 2.7).

## 2.4. Operating Model

The OM is a discrete-time, individual-based model (IBM) implemented in R based on the 'fishSim' 
package [@R-fishSim]. We modified the package functions to accommodate 
our study needs (e.g., by adding the ability to model protogynous hermaphrodite 
life histories) and rewrote the slower-running R functions in C++ using the 'Rcpp' 
package [@eddelbuettel2013seamless] to increase computational efficiency. The modeled
annual sequence, modeled loosely on Gag life history,
was that sex transition and then spawning occurred at the beginning
of a year $t$, followed by mortality, CKMR sampling, and age incrementation (Figure 1).
For computational efficiency, the simulation starts at age 1, given that most age-0 Gag
are not available to the fishery and suffer high rates of natural mortality.

Probability of reproducing in a given year was described by a logistic function, 
and the sex transition function was modeled as a cumulative normal distribution function 
(Figure S1). In order to restrict who can mate with whom in any given year, we assigned
each individual as belonging to one of 40 implicitly spatially-distinct breeding groups. 
Within the spawning season, mating was random and restricted to individuals within the same 
breeding group. All offspring were female 
at birth and entered the population at age 1 as members of their parents’ breeding 
group. Females were allowed to switch breeding groups (see Supplement for details). Males 
did not switch breeding groups but instead remained in the breeding group of their 
sex transition year until death, consistent with the view that male Gag stay on the 
spawning grounds year-round while females migrate between spawning and feeding areas.

Simulated reproductive value for females and males of age $a$, denoted as  $\mathrm{RV}_{a,F}$ and
$\mathrm{RV}_{a,M}$, respectively, were used to produce age-1 offspring and assign parentage.
$\mathrm{RV}_{a,F}$ represents the age-specific expected number 
of age-1 offspring produced per female. This was modeled as a power function 
(Figure 2A) derived 
from female Gag batch fecundity-at-age data, and scaled to achieve a stable population in 
the absence of fishing. To simulate recruitment variability, $\mathrm{RV}_{a,F}$ 
was randomly scaled up or down annually by sampling a reproductive value, 
$\mathrm{RV}_{a,t,F,b}$ for each breeding group, $b$, from a $\mathrm{Uniform}(0,1)$ 
distribution and rescaling such that $\sum_a{\mathrm{RV}_{a,F}e^{\eta_t}} = \sum_a{\sum_b\mathrm{RV}_{a,t,F,b}}$.
Here, $\eta_t$ represents the annual scaling factor, ensuring that the total RV 
across all ages and breeding groups remains consistent.
The realized number of age-1 offspring in a given year $t$
for a given female of age $a$ in breeding group $b$ was a Poisson random draw from 
$\mathrm{RV}_{a,t,F,b}$. Thus, all females of the same age, in the same year, and in the same breeding group
had the same probability of reproducing. 

All males were considered mature, and $\mathrm{RV}_{a,M}$ 
was used as the age-specific probability of being the father
to any age-1 offspring produced in the male's breeding group. The true shape and 
parameter values of $\mathrm{RV}_{a,M}$ are unknown. Here, we modeled it as a power function, and in 
the base OM its exponent was set equal to that of $\mathrm{RV}_{a,F}$ (Figure 2A), 
which corresponds to assuming $\mathrm{RV}_{a,M}$ equals $\mathrm{RV}_{a,F}$ (the current
assumption of the Gag stock assessment). Mating was assumed to 
be fully random within breeding groups. 

Gag life history data and age schedules, including sex change, (Figure S1) were 
taken from the 2022 Gag stock assessment [@SEDAR2022]. Sex change probability was assumed to be 
invariant because there is no strong evidence for sex change plasticity in Gag. We note 
here that simulation studies modeled on species that exhibit phenotypic plasticity in sex change 
should allow females to transition at younger ages 
as males become scarce in the population.

To simulate total
annual survival probability at age, we followed the fisheries modeling approach 
of separating survival into age-specific, instantaneous natural
and fishing mortality components, $M$ and $F$, respectively.
Fishing mortality was further separated by fishing fleet. Traditional fisheries models
use stock-recruitment relationships to represent the density-dependent effects that
result in reduced recruitment at low spawner abundance. We instead explicitly incorporated
density-dependent age-1 survival into our simulations, such that
age-1 natural mortality had density-independent as well 
as density-dependent components. The density-dependent component was assumed to 
scale linearly with juvenile (ages 1 and 2) relative densities (relative to unfished time, $t_0$) such that 
at a relative juvenile density of 1, age-1 density-dependent survival was about 
17 percent, and at a relative juvenile density of 0.1 it was about 84 percent. 
Furthermore, age-1 natural mortality was subject to random variation $\delta$ to 
help produce random recruitment variability. The total probability of survival in year
$t$ for age $a$ was calculated as:
\begin{equation}
S_{a,t} =
\begin{cases}
\begin{aligned}
&\exp\biggl[-\biggl(M_a + \max(1.75 \times \frac{\sum{N_{a\in\{1,2\},t}}}{\sum{N_{a\in\{1,2\},t_0}}}  + \delta_t, 0) + \sum_{f} v_{a,f}F_{t,f}\biggr)\biggr] & \text{if } a = 1 \\
&\exp\biggl[-\biggl(M_a + \sum_{f} v_{a,f}F_{t,f}\biggr)\biggr] & \text{if } a > 1
\end{aligned}
\end{cases}
\end{equation}
where $F_{t,f}$ is the mortality due to fleet $f$ and $v_{a,f}$ is the
vulnerability of age-$a$ fish to fleet $f$. While in the real-world Gag fishery, there are
multiple recreational and commercial hook-and-line fleets, we simulated a 
single recreational and a single commercial fishery for simplicity.
Vulnerability (i.e., the combined effect of gear selectivity 
and fish availability) for the recreational fleet was modeled as a dome-shaped 
gamma function while commercial vulnerability was specified as a logistic function 
(Figure S2A), meaning the oldest individuals are only selected 
for by the commercial fleet. The simulated fishing mortality time series (Figure 
S2B) followed the trend estimated by the 2022 Gag
stock assessment, which was intended to result in overfishing and stock
depletion, mimicking the fate of the real Gag population. 

We used random draws from a uniform distribution to decide if an individual died, 
transitioned or reproduced (females) in a given year.  If an individual’s specific 
age-based probability was higher than the random draw, the event occurred. For any 
fish that died, the cause (i.e., natural mortality, recreational fishing or commercial 
fishing)  was determined based on the  normalized instantaneous rates for that age class and year.

The simulation was run for 119 years, starting with an unfished population and 
ending with a highly depleted population. For each OM configuration, 40
independent replicates of the population were simulated, providing 40
unique whole-population pedigrees.

## 2.5. Sampling Model
Samples for CKMR  (i.e., age 1 for juveniles and ages 4+ for adults) were drawn
during the final three years of the simulation, from all individuals that died from fishing mortality.
The kin relationship between every pair of sampled individuals was identified from the pedigree, 
and mtDNA identity by descent was determined based on the sex of the parent (for POPs) 
or the shared parent (for HSPs).
Pairs that did not share mtDNA identical-by-descent were assigned matching
mtDNA with probability $1-H_\mathrm{mt}$. These observations were tallied according
to both individuals' birth years as well as the first-born's sampling year and sex at time of sampling.
Only pairs of samples with different birth years ("cross-cohort
comparisons") were included in the CKMR likelihood to minimize the effect of
interannual and inter-individual variation in the number of offspring [@waples2022close],
but all samples were used for both HSPs and POPs. The total number of CKMR samples collected across 
all sampling years was based on the
$10\sqrt{N}$ heuristic from @bravington2016close.  We note that this heuristic was developed in the context of 
using POPs for CKMR and that considerably more refined approaches for the design of CKMR experiments are 
available; nonetheless, this heuristic could be easily applied to thousands of simulated data sets to ensure 
that sampling effort was appropriate for the simulated abundance (see Supplemental Material
for an evaluation of different sample sizes and a longer sampling time frame). We used the average simulated abundance of
age 2+ individuals over the CKMR sampling period as $N$ in the total sample
size calculation and the average sex ratio using all ages to partition samples between males and females.
Thus, the sample sizes varied across simulation replicates of the OM. 
The effect of increasing or decreasing the relative proportion of males sampled is explored in the supplement.
The total genetic samples
were spread out over the CKMR sampling period, such that if there were fewer
sampling years, more samples per year would be collected. For each of the 40
simulated pedigrees per OM configuration, dead individuals were subsampled with a different random seed ten 
times to create ten distinct CKMR data sets per pedigree, for a total of 400 CKMR data sets per
OM configuration (the subsampling effect is explored in the Supplementary Material).

The simulated fisheries data consisted of a relative recruitment index and age-composition data, 
both generated with negligible sampling error. These data were made essentially deterministic to isolate 
uncertainty arising from the CKMR and hermaphroditic components of the model, without confounding effects 
from traditional fisheries data uncertainty. Catch data were not generated because the absolute scale of 
landings and discards is highly uncertain for protogynous 
species in the southeastern U.S. and Caribbean. The assessment model thus had to rely on the CKMR data 
alone to estimate scale (i.e., total abundance).

## 2.6. Estimation Model

We employed a standard age-structured deterministic fishery population dynamics
model and integrated the CKMR estimators derived in Appendix A to see how well the 
simulated sex-specific abundances could be estimated, given the data generated from the OM. 
All life history and fishery vulnerability parameters except $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ 
and $\mathrm{RV}_{a,M}$ were fixed at the true value (i.e., assumed known). 
The inflection point of the two-parameter 
cumulative normal distribution $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
function was always assumed known but the standard
deviation, $\mathrm{STF}_{SD}$, which controls the function's shape was estimated in some EM runs. 
The exponent of the $\mathrm{RV}$
power function was assumed known in the EM only for females 
($\mathrm{RV}_F^\mathrm{OM} = \mathrm{RV}_F^\mathrm{EM}$, more on this in Section 2.7). 
Additional estimated parameters were mean
recruitment $\rbar$ and fleet-specific fishing mortality rates $F$ as well as annual
deviations for each. No stock-recruitment relationship was simulated or attempted
to be estimated, and unfished abundance was also not estimated since the focus of
the study was to estimate scale (i.e., absolute abundance) and sex ratio in recent years. 
The estimation model was fit to age compositions, the recruitment index, and the CKMR data.
Full equations
for the total objective function and its components are presented in Appendix B. 
The estimation model was implemented in AD Model Builder (ADMB; @fournier2012ad).

## 2.7. Scenarios

We define two scenario groups for evaluating the performance of the CKMR
estimators derived in Appendix A. The first is focused on the effect of parental sex 
assignment misspecification and the second explores uncertainty surrounding male contribution to reproductive success.
The sex of the shared parent in HSPs is not observed so mtDNA can be used to infer it.
In sex-changing species, the sex of the parent at the offspring's time 
of birth in a POP might also be uncertain, necessitating the use of mtDNA for 
POPs as well.  (Equations A8 and A14-A16, Supplemental Information on assumption on 
equifrequent haplotypes). Thus, $\Hmt$ plays an important role in both HSP and POP
calculations in sex-changing species, and two of our scenarios explored the consequences of 
misspecified this quantity in the EM. Another important consideration for protogynous
hermaphrodites is the reproductive value of males. The true $\mathrm{RV}_{a,M}$ in 
protogynous hermaphrodites, including Gag, is unknown; while  sperm limitation is 
hypothesized for Gag, it has not been measured. We, therefore, simulated four alternative 
versions of ‘reality’ in our OM to explore the effect of uncertainty about male contribution to 
reproductive success. The six scenarios are defined as follows and summarized in Table 2:

1. Base (no $\Hmt$ misspecification, no sperm limitation, male RV is equal to female RV):
In the base scenario, $\Hmt$ in both EM and OM ($\Hmte$ and $\Hmts$, respectively)
was set to 1, meaning perfect accuracy in sex assignment of the parent. Furthermore, 
declining male abundance had no effect on productivity (all eggs produced by females will get fertilized, 
regardless of sex ratio) and the male and female (age-specific) reproductive value functions 
($\mathrm{RV}_{a,M}$ and $\mathrm{RV}_{a,F}$, respectively) were equal in both the OM and EM, 
consistent with current Gag stock assessment model assumptions.

2. $\Hmts$ < $\Hmte$: In this scenario, $\Hmts$ was set to 0.9 while $\Hmte$ was set 
to 1,  all other parameters remaining consistent with the base scenario.

3. $\Hmte$ < $\Hmts$: In this scenario, $\Hmts$ was set to 1 while $\Hmte$ was set 
to 0.9,  all other parameters remaining consistent with the base scenario.

4. Sperm Limitation (skipped spawning): Declining male abundance caused an 
increase in the likelihood of female skipped spawning in the OM. The relationship 
between the current proportion of males in a breeding group relative to unfished 
levels, $p_M$, and the probability of females in that breeding group skipping spawning 
was modeled as an exponential function of the form $1-\exp(-{\tau p_M})$ (Figure 2B), where  
$\tau$ controls the slope of the curve. We used a sigmoid function to map probability of 
spawning to age, such that skipped spawning was more likely in younger females. 
Thus, skipped spawning served as a de facto modifier for the age-specific 
probability of a female producing offspring in a given year. We assumed skipped 
spawning was independent between years, meaning there was no systematic skipped 
spawning such as alternate year spawning. No sperm limitation was assumed in the EM.
$\Hmt$ was not misspecified in this or the following two scenarios.

5. Even male RV across ages: $\mathrm{RV}_{a,M}^\mathrm{OM}$ was much flatter 
(nearly age-invariant; Figure 2A) compared to $\mathrm{RV}_{a,F}^\mathrm{OM}$.

6. Steeper male RV with age: $\mathrm{RV}_{a,M}^\mathrm{OM}$ was much steeper
(with older individuals contributing substantially more to reproduction than younger ones; Figure 2A) 
than $\mathrm{RV}_{a,F}^\mathrm{OM}$.

Estimation for the six scenarios was performed under four configurations 
of treating $\mathrm{RV}_{a,M}$ and $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ as fixed or estimated. Either both quantities 
were fixed, both were estimated, or one was fixed and the other estimated. This was done to explore the 
information content of the CKMR data and how the model behaves and attempts to compensate for different 
types of misspecification (skipped spawning, wrong $\mathrm{RV}_{a,M}$, wrong $\Hmt$).
Importantly, when $\mathrm{RV}_{a,M}$ 
was fixed, it was always fixed at the value corresponding to female reproductive value ($\mathrm{RV}_{a,F}$), 
mimicking real world stock assessment assumptions.
Consequently, $\mathrm{RV}_{a,M}$ was misspecified in the EM for Scenarios 5 and 6, allowing 
us to explore the consequences of getting $\mathrm{RV}_{a,M}$ wrong in the assessment. 
When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was fixed, on the other hand, it was always fixed at the true value. Allowing 
the model to attempt to estimate $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was done to determine whether the CKMR data 
could allow estimating the sex ratio.

We refer to the 24 combinations of the six scenarios and four estimation configurations as 'cases' for the
remainder of this paper. Each case consisted of 40 simulated populations, each subsampled ten times 
to generate unique CKMR data sets, resulting in 400 iterations per case.

## 2.8.  Model Evaluation
For every assessment year $t$ of iteration $k$ within case $c$, we calculated percent relative error ($\mathrm{PRE}$)
between simulated and estimated sex-specific 
adult abundance and sex ratio as:
\begin{equation}
\mathrm{PRE}_{t,k,c} = \frac{(\mathrm{EV}_{t,k,c}-\mathrm{SV}_{t,k,c})}{\mathrm{SV}_{t,k,c}}\times 100,
\end{equation}
where $\mathrm{EV}$ = estimated value and $\mathrm{SV}$ = simulated value. Sex ratio was
defined as the number of males divided by the total number of 3+ year-olds. 
We calculated the median of $\mathrm{PRE}_{t,k,c}$ to obtain one $\mathrm{PRE_{k,c}}$ value per iteration. 
We chose the median over the mean here after inspecting the PRE values within case (Figures S3 and S4). 
For most cases, PREs were flat over years for all iterations, so median and mean perform similarly. 
There were, however, a number of iterations (only for the males and primarily within two cases; Figure S4) 
where extreme PRE values early in
the time series distort the mean while the median captures the dominant signal.

To characterize accuracy and precision by case, we calculated the median and interquartile range (IQR), respectively, of
$\mathrm{PRE_{k,c}}$. The distributions of $\mathrm{PRE_{k,c}}$ were well-behaved for most of the
cases, but some exhibited right-skew and extreme tails (Figure S5) and the median better describes
the bulk of the distribution for those cases than the mean.

Additional model evaluation metrics, including PRE for main model parameters and normalized 
root mean squared error (NRMSE), which provides a mean-based complement to the median PRE, 
are provided in the Supplemental Material.

# 3. Results

## 3.1. Simulated Population

At the beginning of the simulation, the median number of adults across all OM
pedigrees was about 690,000 adult individuals (ages 3+) and about 27% of them were male.
In the final year, the median abundance was ~230,000 individuals without 
the productivity-loss feedback loop from sperm limitation
and ~100,000 with sperm limitation, and about 3% of them were male. When sperm limitation
was present the population trajectory started diverging from the scenarios without
sperm limitation in the final 35-40 years of the simulation (Figure 3).

## 3.2. Simulated CKMR Samples

The median number of females sampled in the base scenario was about two orders of magnitude higher 
than the number of males sampled (median of 3,110 females vs 28 males; Table S1, Figure S6).
The median number
of HSPs found with different mtDNA ($\HSPD$) was about sixteen times higher (758 vs 47) than
HSPs sharing mtDNA ($\HSPS$). Additionally, POPs with matching mtDNA were 
seven times more common than those with different mtDNA (median of 36 vs 5; Table S1, 
Figures 4 and S7). 

Some differences in the number of pairs found were apparent across scenarios
(Figure 4). First, the number of unrelated pairs was much
smaller for skipped spawning than the other scenarios while
the number of related pairs was not. This is because abundance was lower for this
scenario and so fewer individuals needed to be sampled to obtain similar numbers of
related pairs. Additionally, the number of $\HSPS$
was higher for the skipped spawning scenario because younger females
were more likely to skip spawning than older ones, meaning comparatively fewer
females contributed offspring and the likelihood of finding
HSPs sharing a mother increased.
The next major effect was that of $\mathrm{RV}_{a,M}^\mathrm{OM}$ on $\HSPD$.
When $\mathrm{RV}_{a,M}^\mathrm{OM}$ increased more steeply with age, as would be
expected based on the size advantage model, the number of $\HSPD$ increased about
three-fold compared to the base scenario (Figure 4, Table S1).
The explanation is analogous to finding more $\HSPS$ when there is skipped spawning. 
If $\mathrm{RV}_{a,M}^\mathrm{OM}$ is steeper,
older males (of which there are fewer than younger males) are much more successful at fathering offspring and
hence the chance of finding $\HSPD$ increases.
The opposite is seen when $\mathrm{RV}_{a,M}^\mathrm{OM}$ varies little with age. Here,
we only found about two-thirds the number of $\HSPD$ compared to the base scenario
(Figure 4)
because now younger males are also very successful and the offspring are more
evenly spread out among fathers, reducing the likelihood of encountering HSPs sharing
a father.  

## 3.3. Estimation Model Convergence

The estimation model was fitted to 2,400 different data sets: 40 pedigrees subsampled ten times each
for six scenarios (Table 2). For each of those data sets, estimation was performed
under four parameter estimation combinations of either fixing or estimating 
$P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_{a,M}$ for a total of 9,600 model runs.
All in all, ADMB
failed to converge, after fitting the model up to three times with different data
weights, for  227  runs (2.4%). 
An additional 779 EM runs (8.1%) were excluded because
parameters were estimated at bounds. Most of these were due
to the $\mathrm{RV}_{a,M}^\mathrm{EM}$ function exponent parameter. Finally, 38 additional runs (0.4%) were excluded
because the maximum parameter gradient exceeded 0.1.
Overall, 89% of the simulated data sets produced model fits
that were used in subsequent analyses. Fifteen percent of runs converged and passed
the data weight checks after the first time fitting the model and an additional
28% after the second time. If the model ran 3 times, we used the
run with the lowest negative log likelihood amongst those 3 runs.

## 3.4. Base Scenario Estimation Performance 

In the base scenario ($\Hmte$ = $\Hmts$, no skipped spawning, 
$\mathrm{RV}_{a,M}^\mathrm{OM}$ = $\mathrm{RV}_{a,F}^\mathrm{OM}$), male and female abundance were estimated very well
(Figure 5A, Table S2). Estimates for both sexes were highly accurate
($\mathrm{|PRE_{c}}| < \pm 2\%$) regardless of whether $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ 
and $\mathrm{RV}_{a,M}$ were estimated or
fixed. Precision, however, was higher when both these functions were assumed
known ($\mathrm{IQR_{c}} {\approx} 7\%$). Precision was lowest for males when both functions were estimated
($\mathrm{IQR_{c}} = 33\%$; Figure 5A). 

## 3.5 Effect of Parental Sex Assignment Misspecification

Misspecifying $\Hmt$ in the EM had major effects on abundance estimation
performance, but only when at least one of $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_{a,M}$ 
was estimated rather than assumed known (Figure 5B&C).
When $\Hmts$ < $\Hmte$, female abundance was systematically underestimated 
unless both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_{a,M}$ were
fixed in the EM ($|\mathrm{PRE_{c}}| <62\%$; Figure 5B). Conversely, females were systematically overestimated
when $\Hmte$ < $\Hmts$ and $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and/or $\mathrm{RV}_{a,M}$
were estimated, with $\mathrm{PRE_c}$ values ranging from 117 to 335$\%$ (Table S2) --- the largest
observed in this study. Precision was also quite low for these cases, $\mathrm{IQR_c}$ ranging from
91 to 254% (Table S2). Precision in female abundance estimates for the same parameter estimation options 
for $\Hmts$ < $\Hmte$, on the other hand, was much higher ($\mathrm{IQR_c} < 6\%$).

For males, results were more nuanced. When only $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was assumed known, male
patterns mirrored that of the females (Figure 5B&C)
resulting in accurate sex ratio estimates (Figure 6B&C)
despite inaccurate absolute abundance estimates in the case of estimated $\mathrm{RV}_{a,M}$. This highlights the near-total
control the sex transition function in the EM has over estimated
sex ratios---when it was assumed known, sex ratios were almost always accurate and precise (Figure 6, Table S2). 
When both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_{a,M}$ were estimated,
male abundance estimates were fairly accurate ($|\mathrm{PRE_{c}}| \pm 9\%$), in stark contrast to 
female abundance estimates. When only $\mathrm{RV}_{a,M}$ was estimated but
$P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ remained fixed, male abundance was systematically
inaccurate in the opposite direction to females (Figure 5B&C).
Consequently, allowing
$P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ to be estimated when $\Hmt$ was misspecified in the EM resulted in highly
inaccurate sex ratio estimates (Figure 6B&C).

When $\Hmts < \Hmte$, the observed number of $\HSPS$ was higher than when $\Hmts = \Hmte$
(Figure S8). Thus when the EM sees more $\HSPS$ than it expects
in the case of $\Hmts$ < $\Hmte$,
it attempts to reduce the number of expected females to match what it sees in the genetic
data (Figure 5B). It does this by underestimating mean recruitment (Figure S9)
along with changing the shape of $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_{a,M}^\mathrm{EM}$,
if allowed (Figure 7D).
When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ is fixed, thereby locking in the sex ratio, the EM can only reduce the expected
number of females by also lowering the expected number of males (Figure 5B), and it does this by
flattening the estimated $\mathrm{RV}_{a,M}$ function (Figure 7D). This also reduces
the expected number of $\HSPD$, in contradiction to what the genetic data for the
males are indicating. When allowed, the EM will flatten the estimated $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ (Figure 7D) which
results in females turning into males at a younger age, reducing the shift in $\mathrm{RV}_{a,M}^\mathrm{EM}$
required to reconcile observed and expected numbers. The opposite is true when the EM sees fewer
$\HSPS$ than it expects in the data for the $\Hmte$ < $\Hmts$ scenario: recruitment is now
overestimated (Figure S9) and the estimated
$P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_{a,M}$ functions are steeper to create
more expected females (Figure 7C).

## 3.6. Effect of Uncertainty about Male Contribution to Reproductive Success

### 3.6.1. Skipped spawning
When skipped spawning was simulated but not accounted for in the EM, younger females were more likely to skip spawning
than older ones. Fewer younger females are
contributing offspring, effectively shifting $\mathrm{RV}_{a,F}^\mathrm{OM}$ to the right, towards older ages. 
Given that fewer females contribute to reproduction, the
chance of encountering $\HSPS$ increases against the EM's expectation and female abundance is underestimated (Figure 5D).
Under a known $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$, underestimation for both sexes was
low when $\mathrm{RV}_{a,M}$ was also known ($|\mathrm{PRE_{c}}| < 5\%$) but increased
when $\mathrm{RV}_{a,M}$ was estimated ($|\mathrm{PRE_{c}}| <20\%$; Figure 5D). 
When $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was estimated, it was estimated to be slightly less steep than simulated
(Figure 7B), and the resulting female underestimation increased to 
${|\mathrm{PRE_{c}|\approx} 28\%}$ but male estimation improved ($|\mathrm{PRE_{c}}| <5\%$; Figure 5D).
This is because a less steep $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ would turn more young females into males, thereby increasing
the number of males estimated. Consequently, male sex ratios when skipped spawning was simulated
were overestimated when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was allowed to be estimated
($\mathrm{PRE_{c}} < 39\%$, $\mathrm{IQR_{c}} = 49\%$;
Figure 6D; Table S2). Precision in male and female abundance estimates was fairly high
for this scenario ($\mathrm{IQR_{c}} < 13\%$) except for males when both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
and $\mathrm{RV}_{a,M}$ were estimated ($\mathrm{IQR_{c}} = 35\%$; Table S2).

### 3.6.2. Differing simulated reproductive value between the sexes
When the EM was run with $\mathrm{RV}_{a,M}$ equal to $\mathrm{RV}_{a,F}$ but the OM simulated them to be different
(scenarios 5 and 6), 
male abundance was overestimated when simulated $\mathrm{RV}_{a,M}^\mathrm{OM}$ was flatter
and underestimated when it was steeper than $\mathrm{RV}_{a,F}^\mathrm{OM}$; in both cases,
inaccuracies were lower when $\mathrm{RV}_{a,M}$ was allowed to be estimated 
($|\mathrm{PRE_{c}}|\; 2-18\%$) than when fixed at the wrong value ($|\mathrm{PRE_{c}}|\;50-67\%$; Figure 5E&F).
Female abundance estimates were only inaccurate when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was fixed
while $\mathrm{RV}_{a,M}$ was also fixed at the wrong value ($|\mathrm{PRE_{c}}| \pm58\%$), matching
the direction of male inaccuracy. However, when either $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ 
or $\mathrm{RV}_{a,M}$ was allowed to be
estimated, median female abundance estimates were accurate within $\pm{5}\%$ (Figure 5E&F). 
Precision in male and female estimation was variable between cases, as low as 
$\mathrm{IQR_{c}} = 4\%$ and as high as $\mathrm{IQR_{c}} = 92\%$ (Table S2).
Sex ratios were least accurate when
$P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was allowed to be estimated but $\mathrm{RV}_{a,M}$ continued to be
misspecified ($|\mathrm{PRE_{c}}| \pm65\%$). Allowing $\mathrm{RV}_{a,M}$ to be estimated improved accuracy
to within $\pm{20}\%$ (Figure 6E&F). Precision in sex ratio estimates was 
very low for $\mathrm{RV}_{a,M}^\mathrm{OM}$ steeper than $\mathrm{RV}_{a,F}^\mathrm{OM}$
and estimating both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_{a,M}$ 
($\mathrm{IQR_{c}} = 92\%$; Table S2).

## 3.7. Ability to estimate sex transition and male reproductive value
The sex transition function shape was overall well estimated with little variability,
except when $\mathrm{RV}_{a,M}^\mathrm{OM}$ was steeper than 
$\mathrm{RV}_{a,F}^\mathrm{OM}$ and $\mathrm{RV}_{a,M}$ was also estimated (Figure 7F). 
The $\mathrm{RV}_{a,M}$ function was much more variable (Figures 7 and S10).
Generally, the median estimated $\mathrm{RV}_{a,M}$ shapes matched the simulated curvatures,
with the notable exception of when $\Hmts < \Hmte$ (Figure 7D).
However, in almost all cases, some runs estimated incorrect curvatures (i.e., estimated
$\mathrm{RV}_{a,M}$ increased at an increasing rate with age when it was simulated
to increase at a decreasing rate with age, and *vice versa*).

The standard deviation of $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$
($\mathrm{STF}_{SD}$) and the exponent of $\mathrm{RV}_{a,M}^\mathrm{EM}$ were
strongly positively correlated (Figure S11), meaning as the estimated $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$ function
became flatter, $\mathrm{RV}_{a,M}^\mathrm{EM}$ became steeper. 
$\mathrm{STF}_{SD}$ was moderately negatively correlated with mean recruitment (Figure S11).

# 4. Discussion

This study is the first, to our knowledge, to investigate the application of CKMR to a protogynous hermaphrodite 
in an integrated assessment framework. By using a data-generating model that differed
in structure from the estimation model, we were able to confirm that the CKMR estimators
we developed were correct when the base assumptions of the data-generating model were met.
From that point, one can systematically start violating these assumptions to explore
which of them, in isolation or combined, have the largest impact on estimation performance.

We showed that sex-specific abundance for protogynous hermaphrodites can be estimated in
an integrated stock assessment model and that even key parameters used to calculate expected relative reproductive output, 
such as the shape of the age-specific sex transition function and male reproductive 
value-at-age, are estimable using genetic data, but only as long as there are no 
misspecifications in the
estimation model that bias the expected probabilities of observing CKMR kin pairs. 

## 4.1 Effect of Parental Sex Assignment Misspecification

One misspecification we explored is the level of mitochondrial haplotype diversity in the population,
which describes the probability that two randomly sampled mtDNA sequences are different and in turn
affects our ability to accurately determine the sex of the shared parent of an HSP (or the parent's sex
in the case of a POP where the parent was male at the time of sampling).
To illustrate the importance of correctly specifying $\Hmt$, we focus
here on HSPs. Given that HSPs sharing a
mother are much more rare than HSPs sharing a father under extreme sex ratio skew toward females,
having the wrong information about $\Hmt$ could incorrectly inflate or deflate
the true number of HSPs sharing a mother in the sample, thus producing biased abundance estimates.
In the base configuration, $\Hmt$ in both EM and OM ($\Hmte$ and $\Hmts$, respectively)
was set to 1. In two alternative configurations we first set $\Hmts$ higher than $\Hmte$
(1 vs 0.9) and then *vice versa*. In the case of $\Hmts$ < $\Hmte$, the observed number of HSPs with the same mtDNA ($\HSPS$) 
is larger than the true number of maternal HSPs but the EM won't know this, 
meaning it will see more maternal HSPs than there truly are. This leads to underestimating
the number of females and overestimating the number of males.\footnote{Note that $\HSPS$ is only 
synonymous with maternal HSPs when $\Hmt$ is 1, and $\HSPD$ could be either paternal HSPs or HSPs where
the shared parent changed sex.} Conversely, in the case of $\Hmte$ < $\Hmts$, the observed number 
of $\HSPS$ is closer to the true number of maternal HSPs, 
but the EM discounts the number of maternal HSPs it sees because it 
expects some of them to be HSPs with different mitochondrial DNA ($\HSPD$), and it overestimates female
abundance.

Even though $\Hmt$ can be easily and accurately estimated by typing the mtDNA of
a large sample of individuals,
our results underscore the importance of estimating mtDNA haplotype diversity well in order to be able to
estimate $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$ and $\mathrm{RV}_{a,M}$. Previous studies
have estimated $\Hmt$ for Gag on the West Florida Shelf to be ~0.84-0.98 [@jue2015shelf]. In that study, 
sample size was quite low ($N=83$) and the sequencing targeted between 562 and 781 bp of the D-loop
portion of the mtDNA molecule, with the variation in sequence length and in $\Hmt$ reflecting whether
or not indels were included in the diversity measure. If Gag mtDNA haplotypes are not in fact, as we assumed,
equifrequent, haplotype frequencies will need to be estimated, incorporated into the EM as data, and
accounted for in the likelihood. The implications for CKMR estimators of rare and common haplotypes 
in the population could be explored in future simulation studies.

## 4.2 Effect of Uncertainty about Male Contribution to Reproductive Success

### 4.2.1. Skipped spawning

Gag are unusual in the extreme sex ratio skew they exhibit as a consequence of high exploitation rates. 
Other protogynous hermaphrodites have shown more plasticity in sex change [@warner1982mating; @hamilton2007size], where females start 
transitioning at younger ages as males become scarce. Simulation studies for those species would include 
compensation in sex change in the OM. We did not include such compensation in the OM for this study 
because there is no evidence for a strong compensatory response in sex change for Gag.

The mechanism by which extremely low sex ratios in protogynous hermaphrodites might compromise
reproductive resilience is unknown. In our simulation model, we assumed it would result in some females 
(predominantly the younger ones) having no reproductive success (akin to right-shifting age at maturity), 
while that of most larger, older fish is unchanged. This is analogous to time-varying $\mathrm{RV}_{a,F}$, 
shifting toward a steeper function of age as the sex ratio
declines. Time-varying life history parameters can be included in stock assessment models,
but often the data to allow their estimation are not available.

### 4.2.2. Differing simulated reproductive value between the sexes

Age-specific reproductive contribution is a key quantity determining the outcome of stock assessment
and CKMR models, and our simulations highlight this. For gonochoristic species, the only 
reproduction-related age-schedule deemed to be of importance is female fecundity [@methot2013stock], often calculated as 
maturity times weight-at-age and otherwise based on batch fecundity estimates. For sequential
hermaphrodites, it is common practice to assume that reproductive value scales identically
between males and females, although there is no evidence to support this assumption. In fact,
the size advantage model of sex allocation suggests the opposite [@warner1988sex]. 

In our simulations we
included two scenarios with shapes for male reproductive value at age that differed from that of the
females. We illustrate the expected effect of misspecifying $\mathrm{RV}_{a,M}$ for the case where 
$\mathrm{RV}_{a,M}^\mathrm{OM}$ is steeper than $\mathrm{RV}_{a,F}^\mathrm{OM}$ 
(but $\mathrm{RV}_{a,F}^\mathrm{EM}$ = $\mathrm{RV}_{a,M}^\mathrm{EM}$). Here, younger males in 
the simulated population contribute fewer offspring, and the older males (of which there are
fewer) contribute more to reproduction. Consequently, more HSPs sharing a father are found than what the EM expects. 
This signals to the EM that there are
fewer males in the population and male abundance is underestimated. When $\mathrm{RV}_{a,M}^\mathrm{OM}$
is flatter (i.e., more age-invariant) than $\mathrm{RV}_{a,F}^\mathrm{OM}$ 
(but $\mathrm{RV}_{a,F}^\mathrm{EM}$ = $\mathrm{RV}_{a,M}^\mathrm{EM}$), the opposite is the case: younger males
contribute almost as many offspring as older males, and so the probability of finding HSPs
sharing a father decreases, causing an overestimation of male abundance.  

When the EM was run assuming $\mathrm{RV}_{a,M} = \mathrm{RV}_{a,F}$ but the OM 
simulated $\mathrm{RV}_{a,M} \neq \mathrm{RV}_{a,F}$, this resulted in significant estimation bias. However, 
there was enough information in the data to allow $\mathrm{RV}_{a,M}$ to be estimated with no or little bias
when $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$ was known. These are promising findings 
because $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$ can often be estimated outside the 
stock assessment model [@SEDAR2022] from observed  age-specific proportions of males
and females, even though those estimates are also uncertain [@lowerre2021gag]. 
By contrast, it would be difficult to measure $\mathrm{RV}_{a,M}$ directly. Theoretically,
it can be estimated by dividing the sample age composition of adult males by the sample age composition of 
male parents, which allows separating $\mathrm{RV}$ from selectivity-at-age [@davies2017advice]. However, this 
would require representative sampling and adequate numbers of paternal POPs in the sample.

## 4.3 Limitations

Our simulation model, although fairly complex, did not account for all factors believed 
to be important to population dynamics of exploited protogynous species. The most important of these is likely spatial effects. 
@conn2020robustness showed that CKMR estimators can be biased when there is limited dispersal and 
variation in spatial sampling probabilities. Assessing biases in spatial sampling in marine fish is
challenging given that drivers of spawning site selection, fidelity, and dispersal are rarely known. 
Gag, for example, have discrete offshore spawning sites, with assumed 
male site fidelity, pelagic early life dispersal, and estuarine nursery grounds. Early life dispersal pathways 
vary temporally, being driven by loop-current-associated remote forcing. This affects 
recruitment success [@weisberg2014gag] and likely leads to spatial clustering of related individuals in
estuarine nursery sites. Excluding same-cohort comparisons could help mitigate this limited mixing of juveniles in nursery areas,
but the extent of this effect depends on the degree of year-to-year variability in dispersal pathways, which
is currently unknown. If CKMR sampling is fishery-dependent, the
various fishing fleets would need to be sampled appropriately in order to counter the spatial biases
in fishing and not violate the CKMR assumption of equal sampling probability for all individuals.
Future simulations might explore more fully the effects of spatio-temporal sampling variability and 
age-specific sampling selectivity on estimation performance. 

An additional factor our simulation model did not consider but could be included in future studies 
is variability of size-at-age and consequent size-based effects on individual reproductive output for 
individuals of the same age. Furthermore, we assumed perfect
fisheries data, with a highly accurate and precise index of juvenile abundance and perfect, though
combined-sex, composition data. In reality, abundance indices and age composition data are often imperfect, as is the case for
Gag. The age-0 Gag index is composed of multiple data streams 
whose spatio-temporal footprints have shifted through time, making the index prone to bias and uncertainty
[@ingram2021juvenile]. Gag age composition data are highly uncertain and particularly sparse for the private
recreational fleet [@SEDAR2022], which comprises the majority of the catch. Fully operational simulation studies 
would typically include stochasticity in all sampled data, and the effect of uncertainty in fisheries data 
when combined with CKMR should be examined in future simulation studies. 

We made simplifying assumptions for our CKMR estimators that will need to be addressed in any
practical applications, specifically the ability to age sampled fish without error and to perfectly 
distinguish HSPs from other kinship types. Aging imprecision is expected to increase uncertainty
within assessments, and known age-error matrices can be incorporated and integrated over in 
the CKMR likelihood [@fisch2025expected]. Very dense genetic data that permit the inference of 
segments of identity by descent between pairs of individuals could be used to distinguish 
relationships like HSPs from grandparent-grandchild or full avuncular pairs 
[@browning1998relationship; @kruijver2025upper]. 
However, with marker sets of 1000 to 3000 single nucleotide polymorphisms 
commonly applied in CKMR today, it is impossible to genetically distinguish between HSPs and
either grandparent-grandchild or full avuncular (uncle/aunt-neice/nephew) pairs, because pairs with these
relationships all are expected to share the same fraction of genome with 0, 1, or 2 gene copies
identical by descent [@thompson1975estimation]. Furthermore it can be difficult
to distinguish HSPs from less-related kin types such as half-avuncular. Although it is  
possible to account for the presence of such non-target kin pairs in practice, e.g., by
calculating and including kinship probabilities for grandparents and avuncular pairs in the CKMR likelihood 
[@forland2025close] and/or setting one's false negative rate to ensure distinction between HSPs and half-avunculars,
we did not model such processes in this study.

Additionally, neither our OM nor our EM modeled the type of size variation at age present
in real populations.This matters because many
fish life history traits, including reproductive output [@barneche2018fish; @trippel1997effects],
are more strongly correlated with size than age. Variation in size at age may 
lead to a distribution amongst same-age individuals in reproductive output, and hence in expected relative 
reproductive output (ERRO). If these individual differences persist over time, then 
individuals with higher than average ERRO at one age will have higher than average 
ERRO at a later age.  Such persistent individual differences can inflate the probability 
of encountering HSPs [@waples2022close], and would have to be accounted for explicitly in practical applications.

Our estimation model was, by necessity, much simpler than the assessment model used for Gag in reality. 
Any real-world model would likely require
hand-tuning that could not be implemented in a simulation framework attempting to fit thousands of
models. It remains to be seen whether $P_a^{\mathrm{F}\rightarrow\mathrm{M}}$ and $\mathrm{RV}_{a,M}$ could actually
be estimated in the real world assessment model that includes many more parameters and data sources
than our simplistic model, but perhaps with the right data, this is an achievable goal. 

## 4.4 Potential Applications

Unlike most real world fisheries stock assessments, we did not fit our assessment model to catch data. This
was done because we were interested to see how well the CKMR data alone would do in estimating scale
and because catch data for Gag are highly uncertain, mainly due to uncertainty in estimating recreational
catches which make up the majority of Gag removals in the GOM. It is conceivable that the CKMR data
and catch data provide conflicting information about scale and trends, similar to what is often seen
for composition and index data, that would complicate model fitting; this scenario could be explored
in future simulations. We note that many species for which management advice is needed in the
southestern U.S. and Caribbean are data-poor and for some [e.g., Goliath grouper, *Epinephelus itajara*, 
@porch2006catch], catch data are missing entirely. For these, the prospect of incorporating CKMR into catch-free assessment models is highly promising. The costs of conducting genetic analyses have declined significantly in recent 
years and will likely continue to drop, making routine genetic data collection from both 
fishery-dependent and fishery-independent sources an increasingly attractive option. Ultimately, 
cost-benefit analyses of incorporating genetic data into stock assessments will need to be conducted 
on a case-by-case basis, taking into account required sample sizes, sequencing costs, and expected gains, 
such as improvements in assessment performance.


\newpage

## Acknowledgements

We sincerely thank Paul Conn for his thoughtful and thorough review, 
which greatly improved this manuscript. We would like to thank Rob Ahrens whose
ADMB model code provided as part of his excellent fishery stock assessment class
served as the basis for the assessment model developed in this paper. We also
thank four anonymous reviewers for their helpful suggestions and constructive
feedback that strengthened the manuscript.
A portion of the writing and code development for this paper occurred while
CF and ECA were scientists-in-residence with Matthew G. DeSaix at the
mobile High Altitude Venue for Ecological Analysis, Genetics and
Statistics on location in Moab, Utah
during April 2023.  This is contribution mHAVEAGAS-004.  

## Competing interests

The authors declare there are no competing interests.

## Author contributions

C.F.: Conceptualization, methodology, formal analysis, validation, visualization, writing - original draft, writing - review & editing; 
E.C.A.: Conceptualization, methodology, formal analysis, validation, visualization, writing - original draft, writing - review & editing;
L.E.A.: Resources, methodology, writing - review & editing;
D.S.P.: Conceptualization, writing - review & editing;
S.K.L.B.: Conceptualization, methodology, supervision, funding acquisition, project administration, writing - original draft, writing - review & editing

## Funding

This research was supported by the National Oceanic and Atmospheric Administration 
(NOAA; MARFIN grant numbers NA21NMF4330504 and NA19NMF4540066)
and the U.S. Fish and Wildlife Service Sport Fish Restoration Program 
(Grant numbers F-22AF00983_00, F23AF00721-00, and F24AF00716-00).

## Data availability

All code associated with analysis is available on https://github.com/cfriess/protoherm_ckmr.

\newpage

# References

::: {#refs custom-style="Bibliography"}
:::


\newpage

## Tables



```{r table1, echo=FALSE, warning=FALSE}
kableExtra::kable(table1, 
                  escape = FALSE, 
                  booktabs = TRUE, 
                  caption = "Notation and definitions used in this study", 
                  col.names = NULL,
                  longtable = TRUE) %>%
  column_spec(1, width = "6em") %>%
  column_spec(2, width = "33em") %>%
  #row_spec(0, bold=TRUE) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), 
                position = "left", repeat_header_method = "replace") 

```

```{r table2, echo=FALSE, warning=FALSE}
kableExtra::kable(table2_df,
      escape = FALSE,
      col.names = c("Scenario Group", "Scenario", "Skip. Spawn.", "$\\mathrm{RV}_M$", "$H_{mt}^{\\mathrm{EM}}$"),
      align = c("l", "l", "c", "c", "c"),
      caption = table2_caption,
      booktabs = TRUE, linesep = "") %>%
  kable_styling(full_width = FALSE) %>%
  add_header_above(c(" " = 2, "OM" = 2, "EM" = 1)) %>%
  column_spec(1, bold = FALSE) %>%
  row_spec(1, hline_after = TRUE) %>%
  row_spec(3, hline_after = TRUE)

```


\clearpage
\newpage

# Appendix A. Full CKMR Model Specification

\renewcommand{\theequation}{A\arabic{equation}} \setcounter{equation}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{table}{0}

To specify the CKMR model, we first derive kinship probabilities representing the 
likelihood that any two individuals in a genetic sample are HSPs or POPs. Then we combine those into
two different CKMR pseudolikelihoods that we tested. See Table A1 for a 
summary of notation used in the appendices.


### HSP Kinship Probabilities

With protogynous
hermaphrodites, there are three types of HSPs possible, determined by
the sex of the shared parent when each of the half-siblings was born:

- $F\rightarrow F$:  the shared parent was a female ($F$) when producing the first-born and the second-born
members of the HSP.
- $F\rightarrow M$: the shared parent was a female when producing the first-born member and a male ($M$)
when producing the second-born member of the HSP.
- $M\rightarrow M$: The shared parent was a male when producing both the first- and second-born members
of the HSP.  


We assume that ages of sampled individuals are known without error, sampling is conducted on dead
individuals, and age-specific survival
is independent of sex. We use an age-structured
model with intervals indexed by year. We furthermore assume the species is monandric (i.e., all males
originate from females) and that sex change happens at the
beginning of the interval, followed by reproduction, then mortality/fishing/sampling. Thus,
a fish that transitions to male at time $t$ will also be a male when producing offspring
at time $t$ and when sampled in the fishery at time $t$.  The formulas we present here would need
to be adjusted if these events take place in a different order. Let $i$
denote the first-born and $j$ the second-born of a sampled pair, where $c_i \neq c_j$ (cross-cohort comparisons only).  
The half-sibling probabilities for a protogynous
species must include not only the probability of a common parent surviving from the time $i$ was born, $c_i$, to
the time $j$ was born, $c_j$, but also the probability of a female transitioning to male between $c_i$ and $c_j$. 
Define the age difference between $i$ and $j$ as $d=c_{j}-c_{i}$, denote
by $a^*$ the shared
parent’s age at the time $j$ was born, and define $\ERRO_{a^*,t,g}$ as the expected reproductive
output of all the age $a^*$ individuals of sex $g$ ($g \in \{F, M\}$) at time $t$ relative to the population (Table \@ref(tab:table1)):
\begin{equation}
\ERRO_{a^*,t,g} = \frac{N_{a^*,t,g}^*\Mat_{a^*,g}\mathrm{RV}_{a^*,g}}{\sum_{a\in A} N_{a,t,g}^*\Mat_{a,g}\mathrm{RV}_{a,g}},
(\#eq:erro)
\end{equation}
where $\mathrm{Mat}_{a,g}$ and $\mathrm{RV}_{a,g}$ are maturity and reproductive value, respectively, 
of age $a$ and sex $g$ individuals, and $N_{a,t,g}^*$, is the relative abundance of age $a$ and sex $g$ individuals at time $t$.
We define reproductive value at age as the age-specific relative potential to produce offspring. It is typically
associated with mean female fecundity-at-age (more on this below). For males it would correlate with fertilization rates. 
Conditional on $c_i$, $c_j$, and $a^*$, the probability that a randomly sampled pair is an HSP produced by the $F\rightarrow F$ scenario is
\begin{equation}
\begin{aligned}
P(\mathrm{HSP}_{F\rightarrow F}|a^*,c_i, c_j) &= \ERRO_{a^*-d,c_i,F} \\
 &\times \prod_{b=0}^{d-1} S_{a^* - d + b,c_i + b} \\
 &\times \prod_{a = a^*-d+1}^{a^*} (1 - \tprob_{a}) \\
 &\times \frac{1}{N_{a^*,c_j,F}},
\end{aligned}
\end{equation}
where $S_{a,t}$ is the annual age-specific survival probability, $N_{a,t,g}$ is the total abundance of age $a$ and sex $g$ fish at time $t$,
and $\tprob_{a}$ is the probability that an individual that starts age $a$ as a female switches sex to be
an age-$a$ male ($b$ is just a product index).

The above probability arises because, in order for a pair $i$ and $j$ with a mother of age $a^*$ when $j$ was born to form an  $\mathrm{HSP}_{F\rightarrow F}$, the following
four independent events must occur:

1. $i$ must be born to a female of age $a^* - d$. This occurs with probability $\ERRO_{a^*-d,c_i,F}$.
2. The mother of $i$ must survive from year $c_i$ until the time of $j$'s birth, which
happens according to the age-specific survival rates with probability $\prod_{b=0}^{d-1} S_{a^* - d + b,c_i + b}$.
3. The mother of $i$ must remain a female from the time of reproduction at age $a^* -d$ until the time of reproduction at age $a^*$, which
occurs with probability $\prod_{a = a^*-d+1}^{a^*} (1 - \tprob_{a})$.
4. The mother of $j$ must be the same female as the mother of $i$.  Under the
assumption of equality of the expected individual reproductive outputs for fish of the same
sex and age, this probability is simply the reciprocal of the number of
age $a^*$ females at time $c_j$, which can be written as the reciprocal of the
total female abundance at age $a^*$ at time $c_j$: $1/(N_{a^*,c_j,F})$.


It is important to note that $P(\mathrm{HSP}_{F\rightarrow F}|a^*,c_i, c_j)$ is a
probability conditioned on the mother's
age being $a^*$. We obtain the required marginal probability,
$P(\mathrm{HSP}_{F\rightarrow F}|c_i, c_j)$, by summing the conditional
probability over all values of $a^*$, weighted by the probability that $j$'s mother is
of age $a^*$, which is simply $\ERRO_{a^*,c_j,F}$. 
Thus, the marginal probability is:
\begin{equation}
P(\mathrm{HSP}_{F\rightarrow F}|c_i, c_j) = \sum_{d < a^* \leq A} P(\mathrm{HSP}_{F\rightarrow F}|a^*,c_i, c_j)\ERRO_{a^*,c_j,F},
\end{equation}
where $A$ is the maximum age. Note that, in practice, pairs with $c_i=c_j$ are typically excluded, since half siblings from
the same cohort may have correlated juvenile survival which
could lead to downwardly biased estimates of abundance [@waples2022close].


With that first scenario, $F\rightarrow F$, derived, the remaining two follow along similarly,
with differences only in the term describing the probability of changing sex (or not). Thus,
for the scenario $F\rightarrow M$ to occur, the female must transition from being female to
male between times $c_i$ and $c_j$, and the conditional probability is
\begin{equation}
\begin{aligned}
P(\mathrm{HSP}_{F\rightarrow M}|a^*,c_i, c_j) &= \ERRO_{a^*-d,c_i,F} \\
 &\times \prod_{b=0}^{d-1} S_{a^* - d + b,c_i + b} \\
 &\times \biggl[1 - \prod_{a=a^*-d+1}^{a^*} (1 - \tprob_{a})\biggr] \\
 &\times \frac{1}{N_{a^*,c_j,M}}.
\end{aligned}
\end{equation}

The marginal probability of sampling an $\mathrm{HSP}_{F\rightarrow M}$ is then,
\begin{equation}
P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j) = \sum_{d < a^* \leq A} P(\mathrm{HSP}_{F\rightarrow M}|a^*,c_i, c_j)\ERRO_{a^*,c_j,M}.
\end{equation}

To calculate the probability of encountering an HSP arising from the
$M\rightarrow M$ scenario, there are no sex-transition terms because, in protogynous
hermaphrodites, males remain males until death.  Hence, the conditional probability is
\begin{equation}
\begin{aligned}
P(\mathrm{HSP}_{M\rightarrow M}|a^*,c_i, c_j) &= \ERRO_{a^*-d,c_i,M} \\
 &\times \prod_{b=0}^{d-1} S_{a^* - d + b,c_i + b} \\
 &\times \frac{1}{N_{a^*,c_j,M}},
\end{aligned}
\end{equation}
and the marginal probability is
\begin{equation}
P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j) = \sum_{a^*\in A} P(\mathrm{HSP}_{M\rightarrow M}|a^*,c_i, c_j)\ERRO_{a^*,c_j,M},
\end{equation}


Because we do not observe the sex of the shared parent at the time of $i$ and $j$'s birth, we
cannot know with certainty whether a half-sibling pair falls into the $F\rightarrow F$,
$F\rightarrow M$, or $M\rightarrow M$, categories; however we can gain partial information
by sequencing the HSP members' mitochondrial DNA (mtDNA), which is inherited from the female parent.
Let $a$ be the haplotype carried by the first-born and $b$ the haplotype carried
by the second born, noting that they may carry the same haplotype ($a=b$) or
different haplotypes ($a\neq b$). We denote the mtDNA haplotype data in the pair $\mathrm{mt}_{a,b}$. Including the mtDNA as observed data,
we then calculate the joint probability
of the kinship/relationship and mtDNA haplotypes of the pair.
In the absence of sequencing errors or mutation during the most recent meioses,
under the $F\rightarrow F$, scenario, both members of
the HSP will have the same mtDNA sequence because the two mtDNAs are recently
identical by descent. By contrast, in the $F\rightarrow M$ and $M\rightarrow M$ scenarios,
the mtDNA of the two HSPs are not identical by descent and can be regarded as two
randomly sampled mtDNA sequences from the population.  Taking $f_a$ and $f_b$ to be the
relative frequency of mtDNA haplotypes carried by the first-born and second-born,
respectively, we have

\begin{equation}
 P(\mathrm{HSP}, \mathrm{mt}_{a,b}|c_i,c_j) = 
  \begin{cases} 
    \begin{aligned} 
      & f_a P(\mathrm{HSP}_{F\rightarrow F}|c_i, c_j) \\ 
      &\quad + f_a^2 \biggl[P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j)  + P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j)\biggr],
  \end{aligned} & \mathrm{if}~a=b \\
  \\
  f_a f_b\biggl[P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j) + P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j)\biggr], & \mathrm{if}~a \neq b.
  \end{cases}
  \label{eq:hspmtf}
\end{equation}



### POP Kinship Probabilities

For parent-offspring pair (POP) probabilities, we assume that the sex of sampled individuals
can be determined without error and that samples come from the fishery and are
caught after reproduction in any give year. (Note that knowledge about sampled individuals'
sex is not required if only HSPs are used in the likelihood.)
Since we assume sampling is lethal, the sampling year of the first-born individual in a pair,
$i$, (i.e., the potential parent) must be the same or later than the birth year
of the second-born individual, $j$ (the potential offspring) in order for a pair to be a POP.
The sex of the candidate parent must
be accounted for in the kinship probability, and it is also necessary to
account for the possibility that a female parent changes sex to become
a male before being sampled.

We let $s_i$ denote the sampling year of $i$.  If $i$ is sampled in the same
year as $j$'s birth ($s_i = c_j$),
then $i$ could not have changed sex after producing $j$ (recall that we assume
sex-change occurs each year before reproduction), and the probability of a POP
can be calculated by the standard approach. Taking first the case when
$i$ is a female, a pair
$(i,j)$ with $c_i < c_j$ and $s_i = c_j$ will be a
$\mathrm{POP}$ if the following two events occur:


1. $j$ must be born to a female
that is the same age as $i$ at time $c_j$. At time $c_j$,
$i$ will be of age $d = c_j - c_i$, and the probability of being produced by an
age-$d$ female in year $c_j$ is the expected relative reproductive output,
$\ERRO_{d,c_j,F}$.
2. From amongst the $N_{d,c_j,F}$ age-$d$ females producing offspring
at time $c_j$, the mother of $j$ must be just one of them, $i$, which occurs with
probability equal to $1/(N_{d,c_j,F})$.

Thus, we have, if $i$ is a female when sampled at time $c_j$,
\begin{equation}
P(\mathrm{POP}|c_i, c_j, g_i=F,s_i=c_j) = \frac{\ERRO_{d,c_j,F}}{N_{d,c_j,F}},
\end{equation}
where $g_i=F$ indicates that $i$ was of sex $F$ at the time of sampling.  The above
translates easily to the situation when $i$ is a male:
\begin{equation}
P(\mathrm{POP}|c_i, c_j, g_i=M,s_i=c_j) = \frac{\ERRO_{d,c_j,M}}{N_{d,c_j,M}}.
\end{equation}
It is worth noting that, if we write the expected total reproductive output of sex-$g$ individuals
at time $t$ as $\ETRO_{t,g} = \sum_{a\in A} N_{a,t,g}\Mat_{a,g}\mathrm{RV}_{a,g}$ (this is
simply the product of $N_{t,g}$ and the denominator in \@ref(eq:erro)), we can rewrite these probabilities as:
\begin{equation}
P(\mathrm{POP}|c_i, c_j, g_i,s_i=c_j) =  \frac{\mathrm{RV}_{d,g_i} \mathrm{Mat}_{d,g_i}}{\ETRO_{c_j,g_i} }
(\#eq:popsimple)
\end{equation}
for individual $i$ of sex $g_i$, with  $\mathrm{Mat}_{d,M} \equiv 1$ because in protogynous hermaphrodites,
all males are mature.


When sampling of $i$ occurs a year or more after the birth of $j$, it is necessary
to account for the possibility that the parent changed sex after giving birth to
$j$ but before being sampled. If $i$ is a male at time $c_j$, then it will also
be male at all time points in the future.  This POP scenario is denoted by $\mathrm{POP}_{M\rightarrow M}$
and the probability of its occurrence requires no correction due to changing sex, and
thus we can combine this with the result in \@ref(eq:popsimple) for males, giving:
\begin{equation}
P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i, g_{i,s_i} = M) = \mathbbold{1}\{s_i \geq c_j\} \frac{\mathrm{RV}_{d,c_j,M} \mathrm{Mat}_{d,c_j,M}}{\ETRO_{c_j,M} },
\end{equation}
where $\mathbbold{1}\{s_i \geq c_j\}$ is an indicator function that evaluates to 1
if $i$ was sampled after $j$'s birth, and 0 otherwise, and where we added the $s_i$ subscript to $g_i$ to emphasize that $g_{i, s_i}$ is referring to the sex of the
candidate parent at the time of its sampling at $s_i$.

If $i$ was female at time $c_j$, but sampled at time $s_i > c_j$, then there is a
chance that $i$ will have changed sex.  We denote the two possibilities as
$\mathrm{POP}_{F\rightarrow F}$ for the case with no sex change, and $\mathrm{POP}_{F\rightarrow M}$
for the case when sex change occurs. The probabilities for these cases with $s_i > c_j$ are
\begin{equation}
\begin{aligned}
P(\mathrm{POP}_{F\rightarrow F}|c_i, c_j, s_i, g_{i,s_i} = F) &= \mathbbold{1}\{s_i \geq c_j\} \frac{\mathrm{RV}_{d,c_j,F} \mathrm{Mat}_{d,c_j,F}}{\ETRO_{c_j,F}} 
\prod_{d< a \leq s_i - c_i} (1 - \tprob_{a})
\\
P(\mathrm{POP}_{F\rightarrow M}|c_i, c_j, s_i, g_{i,s_i} = M) &= \mathbbold{1}\{s_i \geq c_j\} \frac{\mathrm{RV}_{d,c_j,F} \mathrm{Mat}_{d,c_j,F}}{\ETRO_{c_j,F} } 
\biggl[ 1 - \prod_{d< a \leq s_i - c_i} (1 - \tprob_{a})\biggr].
\end{aligned}
\end{equation}

Of the three above probabilities, only $P(\mathrm{POP}_{F\rightarrow F}|c_i, c_j, s_i)$ can be
used directly when $s_i > c_j$, because we do not observe the sex of $i$ at time $c_j$, and, when
$i$ is a male at time $s_i > c_j$, $i$ could have been either male or female at time $c_j$; however,
as before, we can garner some information from mtDNA.  Thus, similar to the case for HSPs,
conditional on $g_i$,
we calculate the joint probability that the pair
is a POP and that the parent and the offspring carry mtDNA sequence $a$ 
and $b$, respectively.

\begin{equation}
P(\mathrm{POP}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i)=
\begin{cases}
f_a P(\mathrm{POP}_{F\rightarrow F}|c_i, c_j, s_i), & \mathrm{if}~a=b,~g_{i,s_i}=F \\
0,                                                  & \mathrm{if}~a\neq b,~g_{i,s_i}=F \\
\begin{aligned} &f_a P(\mathrm{POP}_{F\rightarrow M}|c_i, c_j, s_i)  \\
&   \quad  + f_a^2 P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i),
\end{aligned} & \mathrm{if}~a=b,~g_{i,s_i}=M \\
 f_a f_b P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i), & \mathrm{if}~a\neq b,~g_{i,s_i}=M
\end{cases}
(\#eq:popcases)
\end{equation}



The probabilities for observing
POPs \@ref(eq:popcases),  or HSPs \@ref(eq:hspmtf),
along with their mtDNA haplotypes 
form the terms in the CKMR pseudolikelihood.



### The CKMR Pseudolikelihood

The pairs of CKMR samples in the data can be summarized into the counts of sample pairs
in different groups according to their covariates: those
born in different years, $c_i < c_j$, with the first-born having sex $g_i \in \{F,M\}$ when
lethally sampled at time $s_i$.   These pairs summarized by $c_i, c_j, g_i, s_i$ may be
further distributed into categories according to their
kin relationship and mtDNA haplotypes. We let $\K{HSP}{c_i,c_j}{g_i,s_i}{a,b}$
$\K{POP}{c_i,c_j}{g_i,s_i}{a,b}$, and $\K{Not}{c_i,c_j}{g_i,s_i}{a,b}$
denote the number of HSPs or POPs or pairs that are neither HSPs or POPs, respectively,
observed among the sample pairs with
a first-born from year $c_i$ sampled as sex $g_i$ in year $s_i$ carrying a mtDNA haplotype
denoted by $a$, and a second-born from $c_j$ carring a mtDNA haplotype denoted $b$. 

Following the standard pseudolikelihood approach for
CKMR (Bravington et al. 2016), we assume that all sample pairs can be treated independently and thus the
joint probability of all the sample counts
for any birth years $c_i < c_j$, sampling years $s_i$, first-born sex at sampling $g_i$, and mtDNA
haplotypes $a$ and $b$ is
approximated by a product of multinomial probabilities. This provides the CKMR likelihood
including both HSPs and POPs:
\begin{equation}
\begin{aligned}
\ell_{HP} \propto \prod_{ {c_i < c_j \atop g_i \in \{F, M\}} \atop Y_\mathrm{lo} \leq s_i \leq Y_\mathrm{lo}}
\prod_{ {a \in \sH \atop b \in \sH}}
& P(\mathrm{HSP}, \mathrm{mt}_{a,b}|c_i,c_j) ^ {\K{HSP}{c_i,c_j}{g_i,s_i}{a,b}} \\
& \times P(\mathrm{POP}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i)^{\K{POP}{c_i,c_j}{g_i,s_i}{a,b}} \\
& \times P(\mathrm{Not}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i)^{\K{Not}{c_i,c_j}{g_i,s_i}{a,b}}
\end{aligned}
\end{equation}
where $Y_\mathrm{lo}$ and $Y_\mathrm{hi}$ are the first and last years, respectively,
when CKMR samples are taken from the fishery, $\sH$ is the set of all known mtDNA haplotypes, and
\begin{equation}
\begin{aligned}
P(\mathrm{Not}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i) = 1 - P(\mathrm{HSP}, \mathrm{mt}_{a,b}|c_i,c_j) - P(\mathrm{POP}, \mathrm{mt}_{a,b} | c_i, c_j, g_i, s_i).
\end{aligned}
\end{equation}
It is worth emphasizing that $\K{POP}{c_i,c_j}{g_i,s_i}{a,b}$ must be
zero for all values of $s_i < c_j$.

We use the log pseudolikelihood $L_{HP} = \log \ell_{HP}$ for inclusion into the EM objective function.


# Appendix B. Estimation Model Objective Function

\renewcommand{\theequation}{B\arabic{equation}} \setcounter{equation}{0}

The joint total objective function, $L_T$, for the observed CKMR and fishery data was the
sum of the weighted CKMR, recruitment index, and age composition components and a 
penalty vector, $L_P$:

\begin{equation}
L_T={\omega_I}L_I + {\omega_C}L_C + {\omega_{HP}}L_{HP} + L_P,
\end{equation}

The likelihood for the index data, $L_I$, was
calculated following the approach described by Walters and Ludwig (1994), by assuming that
the differences between the observed and predicted abundance index
were normally distributed and using the maximum likelihood estimate for survey
catchability $q$. The index log-likelihood (without the constant terms) was thus:

\begin{equation}
L_I = \sum \log(\sigma_I) + (Y_t-\frac{1}{n} \sum Y_t)^2 / 2 \sigma_I^2,
\end{equation}

where $Y_t$ was calculated as:

\begin{equation}
Y_t = log(I_t) - log(\hat{I_t}).
\end{equation}

where $\hat{I_t}$ is the model-predicted age-1 recruitment and $I_t$ is the 
recruitment index.
The composition data likelihood, $L_C$, was specified using a multivariate logistic 
negative log likelihood (Richards et al. 1997). This likelihood, implemented via ADMB's *dmvlogistic* 
function, accounts for differences between observed and expected proportions 
while incorporating overdispersion effects. See Appendix A for the CKMR likelihood, $L_{HP}$.

The penalty vector, $L_P$, consisted of lognormal penalties for the mean fishing mortality
parameters, $\bar{F}_r$ and $\bar{F}_c$, (with $\mu_F$ = -1.2 $\sigma_F$ = 0.8) and the sex
transition function $SD$ parameter, $\mathrm{STF}_{SD}$ 
(with $\mu_{STF}$ = 1.5 $\sigma_{STF}$ = 0.5) 
as well as penalties for recruitment ($P_{\mathrm{Rdev}}$) and fishing mortality ($P_{\mathrm{Fdev}}$) deviations: 

\begin{equation}
L_P = P_{\bar{F}_r} + P_{\bar{F}_c} + P_{STF_{SD}} + P_{\mathrm{Rdev}} + P_{\mathrm{Fdev_r}} + P_{\mathrm{Fdev_c}}.
\end{equation}

The penalties on deviations were set to be weak normal ($\sigma_{dev}$ = 2) until the last 
phase of estimation when they were set as 100 times the sum of the squared elements. 

The initial CKMR and index data weights, $\omega$, were set to one, while the initial composition 
data weight was specified as the number of abundance
index data points (equal to the number of assessment years with data points) 
divided by the age composition data points (equal to number of assessment years times number
of ages modeled, if no missing data). After the initial run, fit statistics were
calculated for each data source and weights were adjusted following the recommendations
of Francis (2011) and Francis (2017). Abundance index weights were
only allowed to increase and CKMR data weights were only allowed to decrease,
while composition data weights could increase or decrease. The model was
refitted up to two times for each iteration. If the model was run more than
once, the run with the highest likelihood was retained for subsequent analysis.


## References

\begin{hangparas}{.25in}{1}

Bravington, M.V., Skaug, H.J., and Anderson, E.C. 2016b. Close-kin mark-recapture.
Stat. Sci. 31(2): 259–274.

Francis, R.C. 2011. Data weighting in statistical fisheries stock assessment models.
Can. J. Fish. Aquat. Sci. $\textbf{68}$(6): 1124–1138.

Francis, R.C. 2017. Revisiting data weighting in fisheries stock assessment models.
Fish. Res. $\textbf{192}$: 5–15.

Richards, L.J., Schnute, J.T., and Olsen, N. 1997. Visualizing catch–age analysis: A case
study. Can. J. Fish. Aquat. Sci. $\textbf{54}$(7): 1646–1658. NRC
Research Press Ottawa, Canada.

Walters, C., and Ludwig, D. 1994. Calculation of Bayes posterior probability distributions
for key population parameters. Can. J. Fish. Aquat. Sci. $\textbf{51}$(3): 713–722.

Waples, R.S., and Feutry, P. 2022. Close-kin methods to estimate census size and effective
population size. Fish Fish. 23(2): 273–293.
\end{hangparas}


```{r tableA1, echo=FALSE, warning=FALSE}
kableExtra::kable(tableA1, 
                  escape = FALSE, 
                  booktabs = TRUE, 
                  caption = "Notation and definitions used in the appendix", 
                  col.names = NULL,
                  longtable = TRUE) %>%
  column_spec(1, width = "6em") %>%
  column_spec(2, width = "33em") %>%
  #row_spec(0, bold=TRUE) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"), 
                position = "left", repeat_header_method = "replace") 

```
