---
title             : "Application of Close-kin mark–recapture (CKMR) to Protogynous Fishes"
subtitle          : "Supplementary Material"
author            : 
  - Claudia Friess, Eric C. Anderson, Lisa E. Ailloud, David S. Portnoy, and
  - Susan K. Lowerre-Barbieri

bibliography      : ["citations_abbrev.bib", "r-references.bib"]
csl               : "elsevier-harvard.csl"

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no
always_allow_html : true

documentclass     : article
output:
  pdf_document:
    keep_tex: true
geometry          : margin=1in
linestretch       : 1.5
fontsize          : 12pt
header-includes   : 
  - \raggedright 
  - \usepackage{color}
  - \usepackage{tipa}
  - \usepackage{wasysym}
  - \usepackage{fancyhdr}
  - \usepackage{mathrsfs}
  - \thispagestyle{empty}
  - \usepackage{hanging}
  - \usepackage{setspace}
  - \usepackage{caption}
  - \usepackage{pdflscape}
  - \setlength{\baselineskip}{1.5\baselineskip}
  - \newcommand{\tprob}{P^{F\rightarrow M}}
  - \newcommand{\wt}[1]{\omega_#1}  % recruitment deviations
  - \newcommand{\rbar}{\bar{R}}  % mean recruitment
  - \newcommand{\fbarr}{{\bar{F}_\mathrm{rec}}}  % mean fishing mortality from rec fishery
  - \newcommand{\frecdev}[1]{fdev^\mathrm{rec}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\fbarc}{{\bar{F}_\mathrm{com}}}  % mean fishing mortality from commercial fishery
  - \newcommand{\fcomdev}[1]{fdev^\mathrm{com}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\Mage}[1]{\dot{M}_#1}  % age-specific instantaneous natural mortality
  - \newcommand{\vulr}[1]{\dot{v}^\mathrm{rec}_#1}  % age-specific vulnerability to recreational fishery
  - \newcommand{\vulc}[1]{\dot{v}^\mathrm{com}_#1}  % age-specific vulnerability to commercial fishery
  - \newcommand{\dFec}[1]{\dot{Fec}_#1}    % age-specific relative fecundity
  - \newcommand{\Af}[1]{\dot{Mat}_#1}    % age-specific fraction of females mature
  - \newcommand{\Pfm}[1]{\dot{P}^{\mathrm{F}\rightarrow\mathrm{M}}_#1}    % age-specific prob of transitioning from female to male
  - \newcommand{\Imp}[1]{\dot{P}^\mathrm{M}_#1}    % equilibrium proportion of males amongst age a fish
  - \newcommand{\It}[1]{\mathring{I}_{#1}}   % recruitment index in year t
  - \newcommand{\pacom}[2]{\mathring{p}^\mathrm{com}_{#1,#2}}   % relative abundance of age a females in year t
  - \newcommand{\pacr}[3]{\mathring{p}^\mathrm{ckmr}_{#1,#2,#3}}   % relative abundance of age a males in year t
  - \newcommand{\Z}[2]{Z_{#1,#2}}   % total instantaneous mortality at time t for age a fish 
  - \newcommand{\AS}[2]{S_{#1,#2}}   % total annual survival
  - \newcommand{\lxf}[1]{s_{#1}}     % survivorship
  - \newcommand{\Nf}[2]{N^\mathrm{fem}_{{#1}, {#2}}}  % number of females of age a at time t
  - \newcommand{\Nm}[2]{N^\mathrm{mal}_{{#1}, {#2}}}  % number of males of age a at time t
  - \newcommand{\K}[4]{\mathring{K}^{\mathrm{#1},{{#4}}}_{{#2},{#3}}}
  - \newcommand{\KHS}[3]{\mathring{K}^{\mathrm{HSP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSs}[2]{\mathring{K}^{\mathrm{HSP},{{#2}}}_{{#1}}}
  - \newcommand{\KPOPT}[3]{\mathring{K}^{\mathrm{POP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSD}[2]{\mathring{K}^{\mathrm{HSP}^{\neq _{a,b}}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with different mtDNA 
  - \newcommand{\KHSS}[2]{\mathring{K}^{\mathrm{HSP}^{=_a}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOP}[3]{\mathring{K}^{\mathrm{POP}^{#3}}_{{#1},{#2}}}  
  - \newcommand{\KPOPF}[2]{\mathring{K}^\mathrm{POP_F}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOPM}[2]{\mathring{K}^\mathrm{POP_M}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA 
  - \newcommand{\HSPD}{\mathrm{HSP}^{\neq}}
  - \newcommand{\HSPS}{\mathrm{HSP}^=}
  - \newcommand{\POPD}{\mathrm{POP}^{\neq}}
  - \newcommand{\POPS}{\mathrm{POP}^=}
  - \newcommand{\KNot}[2]{\mathring{K}^\mathrm{Not}_{{#1},{#2}}}     % number of pairs that are neither HSD or HSS
  - \newcommand{\KNote}[2]{\mathring{K}^{\mathrm{Not}=_a}_{{#1},{#2}}}
  - \newcommand{\KNotu}[2]{\mathring{K}^{\mathrm{Not}{\neq}_{a,b}}_{{#1},{#2}}}   
  - \newcommand{\Hmt}{H_\mathrm{mt}}  % mtDNA haplotype diversity
  - \newcommand{\Hmts}{H_\mathrm{mt}^{\mathrm{OM}}}  % simulated mtDNA haplotype diversity
  - \newcommand{\Hmte}{H_\mathrm{mt}^{\mathrm{EM}}}  % mtDNA haplotype diversity assumed in the EM
  - \newcommand{\Mat}{\mathrm{Mat}}
  - \newcommand{\Fec}{\mathrm{Fec}}
  - \DeclareMathAlphabet{\mathbbold}{U}{bbold}{m}{n}
  - \newcommand{\comment}[1]{\textcolor{red}{#1}}
  - \newcommand{\ERRO}{\mathrm{ERRO}}
  - \newcommand{\ETRO}{\mathrm{ETRO}}
  - \newcommand{\sH}{\mathscr{H}}
---

```{r setup, include = FALSE}
library(tidyverse)
library(stargazer)
library(kableExtra)
library(gridExtra)
library(nplyr)
library(patchwork)
library(ggh4x)
library(cowplot)
library(gridExtra)
library(grid)
library(scales)
library(here)

# caching for quick knitting.  
#    set cache = FALSE to not cache
#    set cache.extra to a new value to remove all caches and re-evaluate
knitr::opts_chunk$set(cache.extra = 4, cache = FALSE) 

source(here("code", "analysis", "results_suppl.R"), local = knitr::knit_global())


```


\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}



## Female Breeding Group Switching in the Operating model

Each year, reproducing females spawned with 
their birth breeding group with probability $\zeta$, interpreted as their spawning 
group fidelity, set to 0.01. Females that switched breeding groups in a given breeding season 
(determined through Bernoulli draws with probability 1 - $\zeta$) switched from their 
birth breeding group $b_0$ to a different group $b_x$ with probability $\psi^{b_0,b_j}$. 
Probabilities $\psi$ were computed as the normalized inverse Euclidean distance 
values of a vector consisting of consistent intervals and having the length of 
the number of distinct breeding groups. This resulted in greater likelihoods of 
moving to a neighboring breeding group than a more distant one. 

## Assumption of Equifrequent Haplotypes in the Simulations

For the purposes of the simulations we present in this paper, we assume that all
mtDNA haplotypes are at the same frequency, (i.e., $f_a = 1/n_H$, for all $a \in \sH$, where
$n_H$ is the number of haplotypes in the set of all mtDNA haplotypes $\sH$). By doing so,
the simulations are greatly
simplified, as it is not necessary to track individual mtDNA haplotypes over generations
across all individuals. Rather, it is sufficient to record whether a sampled pair
shares the same mtDNA sequence or not. Furthermore, this allows us
to summarize the genetic variation in the mtDNA in terms of a single number, 
$\Hmt$, the probability that two mtDNA sequences randomly
sampled from the population are different in sequence.  

When mtDNA haplotypes are assumed equifrequent, we are able to categorize---without
loss of information---sampled pairs not by their specific
haplotypes, $a$ and $b$, but rather by whether
the pair shares the same haplotype or not, regardless of the specific sequence.
Thus in the CKMR pseudolikelihood to analyze the simulations we used joint probabilities
of a pair's type of relationship and whether or not the
members of the pair have the same mtDNA sequence. This joint probability is easily found
by summing the probabilities in (8) and (14).

Summing the first case ($a=b$) of (8) over all haplotypes $a\in\sH$ yields

\begin{equation}
\begin{aligned}
P(\mathrm{HSP}, \mathrm{mt}^=|c_i,c_j) &= P(\mathrm{HSP}_{F\rightarrow F}|c_i, c_j) \\
&\qquad + (1-\Hmt) \biggl[P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j) + P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j)\biggr]
\end{aligned}
\end{equation}
where $\mathrm{mt}^=$ denotes simply that the pair have the same mtDNA.  
Summing the
second case ($a\neq b$) in (8) over all pairs 
$(a,b),\ a\in\sH,\ b\in\sH,\ a\neq b$, gives

\begin{equation}
\begin{aligned}
P(\mathrm{HSP}, \mathrm{mt}^{\neq}|c_i,c_j) &= \Hmt\biggl[P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j) + P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j)\biggr].
\end{aligned}
\end{equation}
where $\mathrm{mt}^{\neq}$ denotes that the haplotypes carried by the pair's members
are of different sequences. 


Similarly, for POPs, summing (14) over different haplotypes provides the
simplification:

\begin{equation}
\begin{aligned}
P(\mathrm{POP}, \mathrm{mt}^= | c_i, c_j, g_i=F, s_i) &=  P(\mathrm{POP}_{F\rightarrow F}|c_i, c_j, s_i)\\
P(\mathrm{POP}, \mathrm{mt}^{\neq} | c_i, c_j, g_i=F, s_i) &= 0 \\
P(\mathrm{POP}, \mathrm{mt}^=|c_i, c_j, g_i=M, s_i) &=  P(\mathrm{POP}_{F\rightarrow M}|c_i, c_j, s_i) \\
& + (1-H_\mathrm{mt}) P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i) \\
P(\mathrm{POP}, \mathrm{mt}^{\neq}|c_i, c_j, g_i=M, s_i) &=  H_\mathrm{mt} P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i) \\
\end{aligned}
\end{equation}

Both (S2) and (S3)) are easily used in a CKMR pseudolikelihood for
pairs categorized according to whether the mtDNA of the members matches or not.

## Additional Performance Metrics

In addition to median Percent Relative Error (PRE), we also calculated Normalized Root Mean 
Squared Error (NRMSE) for every case $c$ as an additional accuracy metric:

\begin{equation}
\mathrm{NRMSE_{c}} = \frac{\sqrt{\frac{1}{n_{c}} \sum_{i=1}^{n_{c}} (\mathrm{SV}_i - \mathrm{EV}_i)^2}}{\frac{1}{n_{c}} \sum_{i=1}^{n_{c}} \mathrm{SV}_i},
\end{equation}

where $n_c$ is the number of simulations within a case. 

The full set of performance metrics for
sex-specific abundance and sex ratio are shown in Table S2. The pattern
of NRMSE was generally similar to median PRE (Figures S12 and S13).Both
metrics identify the problematic scenarios but negative PRE values 
revealed systematic underestimation in some cases that NRMSE alone would not have distinguished 
from overestimation.

## Investigating the Subsampling Effect

We subsampled each of the simulated populations ten times to create multiple data sets per population
for analysis. This was done for computational efficiency (it takes several hours to generate each pedigree),
but it also presents an opportunity to investigate the effect of random chance due to simulated pedigree
as opposed to random chance due to subsampling each pedigree. The subsampling effect was
explored by calculating the intraclass correlation coefficient (ICC) using one-way random effects
models of $\mathrm{PRE_c}$ with simulated population as the grouping factor. 
A separate model was fitted for each sex and each case
for a total of 48 models fitted. A low ICC
indicates a high degree of CKMR sampling stochasticity contribution to total variation and little effect
of population history. Conversely, a high ICC means the subsamples from the same population tend to cluster
together and differ from the subsamples of another population, with some populations consistenly producing
high PREs and others consistently producing lower ones.

We found that ICC was highly variable across cases, ranging from near zero to 0.78 (Figure S14). Without
model misspecification, CKMR sampling stochasticity was the dominant driver of variability (ICC < 0.2). 
ICC became elevated for some model misspecification, particularly when $\mathrm{RV}_{a,M}$ was simulated
to be steeper in the OM than $\mathrm{RV}_{a,F}$ and when $\mathrm{RV}_{a,M}$ was estimated for the
$\Hmte < \Hmts$ scenario (Figure S14). These patterns are also apparent when visualizing PREs
by simulated popuation and case for the females (Figure S15) and males (Figure S16).

## Effect of changing the number of samples collected for CKMR and the timeframe over which they were collected

In order to evaluate the impact of CKMR sample size and sampling time frame, we ran a 
number of supplementary scenarios where the model was configured the same way as for the Base
scenario except for CKMR sample collection. In addition to the $10\sqrt{N}$ sample size
heuristic that served as the base sample size, the model was also run with 50% and 150% of that base sample size. 
We also explored whether 
the time period over which sampling is spread out affects the number of related pairs found and 
EM performance by changing the number of sampling years from three to ten. In the longer sampling scenario,
roughly the same overall CKMR sample size was collected as for the Base scenario, 
it was just spread out over more years (Figure S6).

The number of CKMR pairs found was higher when sampling was conducted
over a shorter time period, and this was most pronounced for
POPs sharing mitochondrial DNA and least pronounced for HSPs and POPs
with different mitochondrial DNA (Table S1, Figure S7). Additionally, the number
of pairs found was about nine times higher in the high (150% $10\sqrt{N}$)  compared to the low (50% $10\sqrt{N}$) and CKMR
sample size runs and about four times higher between the low and
base (100% $10\sqrt{N}$) runs (Table S1, Figure S7).

Estimation accuracy was similar between three and ten years of CKMR sampling (Figure 17).
Accuracy for male estimation was high ($\mathrm{|PRE_{c}}| < \pm 3\%$) for all cases of 
sampling variability. Accuracy results for female
estimation were similar to males, except it was slightly lower for ten years of sampling at 50% base sample size
when the sex transition was estimated ($\mathrm{|PRE_{c}}|< \pm 8\%$; Table S2).
Precision was high across all CKMR sampling scenarios when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and
$\mathrm{RV}_{a,M}$ were known (Figure 17A, Table S2), with $\mathrm{IQR_{c}} < \pm 13\%$.
Precision in abundance estimates increased as CKMR sample size increased and was also slightly
higher for three years of CKMR sampling compared to ten (Figure S17, Table S2).

## Effect of changing the number of males sampled for CKMR

The number of adult males sampled for CKMR in our simulation study was about two orders of
magnitude smaller than the number of females sampled. This might beg the question what would
happen if we were to increase the proportion of males in the sample. We conducted additional
runs on the base and skipped spawning scenarios to explore this question. In one set of runs,
we excluded age 1 females from the sex ratio calculation used to partition the CKMR sample by sex,
which roughly doubled the number of males sampled (Figure S18A) and
in another set, we sampled no males at all. The result was that the number of $\mathrm{POP}^{\neq}$
found roughly doubled when the number of males sampled doubled and decreased to zero when no males
were sampled, but the other kinship types were fairly unaffected (Figure S18B).
For estimation performance, we found that varying the number of males sampled had no effect 
(Figure S19), further emphasizing the minimal impact of
$\mathrm{POPs}$ on our estimator's accuracy and precision. 

## Supplementary Tables


```{r tableS1, echo=FALSE, warning=FALSE, results='asis'}
tS1_cap <- "Median number of adult males and females sampled and number of observed related pairs found by 
scenario as well as a number of supplemental scenarios ('Varied Sampling', where the number of CKMR
samples and sampling time frame were modified) investigating the effect of increasing
the number of years over which CKMR samples are collected (Yrs) and varying the number of 
samples collected (Nsamps; proportion of base sample size, see text for details). The 
number in parentheses is the interquartile range. sHmt = simulated mitochondrial haplotype diversity,
eHmt = mitochondrial haplotype diversity assumed in the EM. mRV = male reproductive value. 
See Table 2 in main text for more information
on the six main scenarios. POD = parent-offspring pairs with different mitochondrial DNA,
POS = parent-offspring pairs with matching mitochondrial DNA, HSD = half-sibling
pairs with different mitochondrial DNA, HSS = half-sibling pairs with matching
mitochondrial DNA."

tab_tex <- knitr::kable(
  samps_table_data, 
  booktabs  = TRUE, 
  caption   = tS1_cap, 
  longtable = TRUE
) %>%
  add_header_above(c(" " = 3, "Number Sampled" = 2, "CKMR Pairs Found" = 5)) %>%
  kable_styling(
    latex_options          = c("hold_position", "repeat_header"),
    position               = "left", 
    font_size              = 10,
    repeat_header_method   = "replace"
  ) %>%
  kableExtra::landscape() %>%
  kableExtra::row_spec(
    0,
    extra_latex = "\\captionsetup{justification=centering, width=\\textwidth}"
  )

# Remove all addlinespace lines
tab_tex <- gsub("\\\\addlinespace\n", "", tab_tex)

cat(tab_tex)

```


```{r tableS2, echo=FALSE, warning=FALSE, results='asis'}
tS2_cap <- "Error metrics for the six main scenarios (24 cases) and the supplementary scenarios where 
CKMR sample size and sampling time frame were varied. PRE = Percent Relative Error. Median PREs were first calculated
across years within model run iteration to obtain a single PRE value per iteration. The PRE and IQR
(interquartile range) values reported in this table represent the median and IQR, respectively, of
those median PREs, by indicated scenario. NRMSE = Normalized Root Mean Square Error. 
(normalized by the mean of simulated values). N represents the number of unique data sets analyzed
for each scenario (400 minus the runs which were excluded due to conversion issues). sHmt = simulated mitochondrial haplotype diversity,
eHmt = mitochondrial haplotype diversity assumed in the EM. mRV = male reproductive value.
See Table 2 in main text for more information on the six main scenarios."
tab_tex <- knitr::kable(error_metrics_full, 
                  booktabs = TRUE, 
                  caption = tS2_cap, 
                  longtable = TRUE,
                  linesep = c("", "", "","", "\\addlinespace") # this doesn't work
                  ) %>%
  add_header_above(c(" " = 3, "Females" = 3, "Males" = 3, "Sex Ratio" = 3)) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"),
                position = "left", font_size = 10,
                repeat_header_method = "replace") %>%
  kableExtra::landscape() %>%
  kableExtra::row_spec(0, extra_latex = "\\captionsetup{justification=centering, width=\\textwidth}")

# Remove all addlinespace lines
tab_tex <- gsub("\\\\addlinespace\n", "", tab_tex)

cat(tab_tex)

```


```{r fig_caps, echo=FALSE}

age_sched_fig_caption <- "Age schedules for natural mortality (dashed line), female probability of reproducing (solid line), 
and sex transition probability (dotted line) used in operating model simulations. These were taken from the gag stock assessment
(SEDAR, 2022).
We allowed the terminal age to act as a plus group, 
rather than killing all individuals that exceed the maximum age (the default behavior of fishSim) but set the natural mortality for this
plus group high to avoid building up a large pool of old individuals."

vulnF_plot_fig_caption <- "Fishing-related parameter values used in the simulations.  A) commercial (solid line) and recreational (dotted line) fishing fleet vulnerability at age and B) commercial (solid line) and recreational (dotted line) fishing mortality."

nsamps_plot_fig_caption <- "Number of adult fish CKMR samples collected by sex for the base OM and EM configuration 
(no skipped spawning, male and female reproductive value are equal, no misspecified $H_{mt}$) for different CKMR sample sizes spread out over three (red) or ten (blue) years of sampling. The percentages refer to multiples of the $10\\sqrt{N}$ sampling heuristic. 
The base scenario collected $10\\sqrt{N}$ samples over three years.
The box represents the
interquartile range (IQR), the line is the median, the whiskers extend to 1.5 IQR and outliers fall beyond that."

supplckmrpairs_plot_fig_caption <- "Number of CKMR pairwise relationships observed for the base OM and EM configuration 
(no skipped spawning, male and female reproductive value are equal, no misspecified $H_{mt}$) for different CKMR sample sizes spread out over three (red) or ten (blue) years of sampling. The percentages refer to multiples of the $10\\sqrt{N}$ sampling heuristic. 
The base scenario collected $10\\sqrt{N}$ samples over three years. 
$\\mathrm{POP}^{\\neq}$ = parent-offspring pairs with different mitochondrial DNA, $\\mathrm{POP}^=$ = parent-offspring pairs with matching mitochondrial DNA, 
$\\mathrm{HSP}^{\\neq}$ = half-sibling pairs with different mitochondrial DNA, $\\mathrm{HSP}^=$ = half-sibling pairs with matching mitochondrial 
DNA. Boxplot elements are as described for Figure S6."


supplckmrpairs2_plot_fig_caption <- "Number of CKMR pairwise relationships by scenario comparing Base to what is observed when
mitochondrial haplotype diversity ($H_{mt}$) is misspecified. $\\mathrm{POP}^{\\neq}$ = parent-offspring pairs with different mitochondrial DNA, $\\mathrm{POP}^=$ = parent-offspring pairs with matching mitochondrial DNA, 
$\\mathrm{HSP}^{\\neq}$ = half-sibling pairs with different mitochondrial DNA, $\\mathrm{HSP}^=$ = half-sibling pairs with matching mitochondrial 
DNA. Boxplot elements are as described for Figure S6."

parnll_plot_fig_caption <- "Estimated parameters for the scenario group exploring the effect of parental sex
assignment misspecification, by whether or not the sex transition function standard deviation parameter, $\\mathrm{STF}_{SD}$ and the exponent of the male reproductive value 
(RV) functions are assumed known (fixed) or estimated (free).
trans = sex transition function, mRV = male reproductive value. Y-axis for mean recruitment and Male RV function exponent truncated to improve readability. Boxplot elements are as described for Figure S6."

par_plot_fig_caption <- "Estimated parameters for the scenario group exploring uncertainty in male contribution to reproductive success
, by whether or not the sex transition function standard deviation parameter, $\\mathrm{STF}_{SD}$ and the exponent of the male reproductive value 
(RV) functions are assumed known (fixed) or estimated (free). Notation is as for Figure S9 and boxplot elements are as described for Figure S6." 

cor_plot_fig_caption <- "Correlation between key estimated parameter pairs (from the ADMB-estimated correlation matrix for
each individual simulation) when both the sex transition function standard deviation and male reproductive value (RV) are
estimated. Fct = function. Boxplot elements are as described for Figure S6."

fac_plot_fig_caption <- "Comparison of accuracy metrics for female abundance estimates by simulated case, including supplemental runs
exploring variation in CKMR sampling.
PRE = Percent Relative Error, NRMSE = Normalized Root Mean Squared Error. RMSE was normalized by dividing by the mean of simulated values."

mac_plot_fig_caption <- "Comparison of accuracy metrics for male abundance estimates by simulated case, including supplemental runs
exploring variation in CKMR sampling.
PRE = Percent Relative Error, NRMSE = Normalized Root Mean Squared Error. RMSE was normalized by dividing by the mean of simulated values."

fpr_plot_fig_caption <- "Percent Relative Error for female abundance estimates for the first 20 of the 40 simulated populations, 
by scenario, and whether the sex transition function (trans) and male reproductive value (mRV) were fixed or estimated. 
Notice
the different scales between panels. One outlier omitted for scenario 'est Hmt < sim Hmt' for better visualization. 
Boxplot elements are as described for Figure S6."

mpr_plot_fig_caption <- "Percent Relative Error for male abundance estimates for the first 20 of the 40 simulated populations, by scenario
scenario, and whether the sex transition function (trans) and male reproductive value (mRV) were fixed or estimated. Notice
the different scales between panels.
Boxplot elements are as described for Figure S6."

sbys_plot1_fig_caption <- "Number of individuals sampled (Panel A) and number of kin pairs found (Panel B) for
the Base and skipped spawning scenarios when varying the proportion of males in the adult CKMR sample. When all
ages are included in calculating true sex ratio used to partition the sample by sex (base male sampling; red), 
fewer males are sampled than when age 1s (who are all female) are excluded from the sex ratio calculation (green). 
Boxplot elements are as described for Figure S6."

sbys_plot2_fig_caption <- "Percent relative error in female (Panel A) and male (Panel B) abundance estimation
for the Base and skipped spawning scenarios when varying the proportion of 
males in the adult CKMR sample. When all
ages are included in calculating true sex ratio used to partition the sample by sex (base male sampling; red), 
fewer males are sampled than when age 1s (who are all female) are excluded from the sex ratio calculation (green).
Boxplot elements are as described for Figure S6."

prebyyearF_plot_fig_caption <- "Percent relative error (PRE) by year for the 24 cases for the females. Scenarios are 
shown in the columns and parameter estimation combination settings in rows. Each line represents one model iteration. 
sHmt = simulated mitochondrial haplotype diversity, eHmt = mitochondrial haplotype diversity assumed in the EM, 
mRV = male reproductive value, trans = sex transition function. The y-axis was truncated for better visualization, 
omitting 20 iterations with PRE values > 500 (6 for eHmt < sHmt, fixed trans \\& est mRV; and 16 for mRV steeper, est trans \\& mRV)"

prebyyearM_plot_fig_caption <- "Percent relative error (PRE) by year for the 24 cases for the males. 
Panel explanation and notation is as in Figure S3. The y-axis was truncated for better visualization, 
omitting 10 iterations with PRE values > 1000, all for the eHmt < sHmt scenario."

predist_plot_fig_caption <- "Distribution of iteration median PREs by case (diferent scenarios
indicated by different colors). For most cases, the distributions were well-behaved,
but some (e.g., mRV steeper, est trans \\& mRV; eHmt < sHmt fixed trans \\& est mRV) 
exhibited right-skew and heavy tails. Notation is as in Figure S3."

icc_plot_fig_caption <- "Intraclass correlation coefficients (ICC) of median PRE by case, 
quantifying the proportion of total variance due to differences among simulated populations 
relative to CKMR sampling stochasticity."

ckmrsamp_plot_fig_caption <- "Percent relative error in estimated male and female abundance time series 
summarized across years, by number of CKMR sampling years (red = 3, blue = 10), CKMR sample size, 
and whether or not sex transition 
($P^{\\mathrm{F}\\rightarrow \\mathrm{M}}$) and male reproductive value 
($\\mathrm{RV}_{M}$) functions are were assumed known (fixed). Results are shown for the base scenario, without
any model misspecification. Y-axis truncated to improve readability. Boxplot elements are as described for Figure S6."

```

## Supplementary Figures

```{r age-sched-plot, echo=FALSE, fig.align='center', fig.height=3, fig.cap = age_sched_fig_caption, fig.width=5, message=FALSE, warning=FALSE}
age_schedule_plot
```

```{r vulnF-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=5, fig.cap = vulnF_plot_fig_caption, message=FALSE, warning=FALSE}
F_vul_plot
```

\newpage
\begin{landscape}
```{r prebyyearF-plot, echo=FALSE, fig.align='center', fig.height=5.5, fig.width=9, fig.cap = prebyyearF_plot_fig_caption, message=FALSE, warning=FALSE}
pre_by_year_females 
```

\newpage

```{r prebyyearM-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=9, fig.cap = prebyyearM_plot_fig_caption, message=FALSE, warning=FALSE}
pre_by_year_males 
```

\newpage

```{r predist-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=9, fig.cap = predist_plot_fig_caption, message=FALSE, warning=FALSE}
grid.arrange(
  arrangeGrob(p1, p2, p3, p4, p5, p6, ncol = 3,
              left = textGrob("Count", rot = 90, vjust = 1),
              bottom = textGrob("Percent Error"))) 
```

\end{landscape}



\newpage
```{r nsamps-plot, echo=FALSE, fig.align='center', fig.height=4.5, fig.width=6.5, fig.cap = nsamps_plot_fig_caption, message=FALSE, warning=FALSE}
nsamps_plot
```

\newpage

```{r supplckmrpairs-plot, echo=FALSE, fig.align='center', fig.height=4.5, fig.width=6.5, fig.cap = supplckmrpairs_plot_fig_caption, message=FALSE, warning=FALSE}
suppl_ckmr_pairs_plot
```

\newpage

```{r supplckmrpairs2-plot, echo=FALSE, fig.align='center', fig.height=4.5, fig.width=6.5, fig.cap = supplckmrpairs2_plot_fig_caption, message=FALSE, warning=FALSE}
suppl_ckmr_pairs_plot2
```

\newpage
\begin{landscape}
```{r parnll-plot, echo=FALSE, fig.align='center', fig.height=5.5, fig.width=9, fig.cap = parnll_plot_fig_caption, message=FALSE, warning=FALSE}
para_res_plot_nll 
```

\newpage

```{r par-plot, echo=FALSE, fig.align='center', fig.height=5.5, fig.width=9, fig.cap = par_plot_fig_caption, message=FALSE, warning=FALSE}
para_res_plot 
```
\end{landscape}
\newpage

```{r cor-plot, echo=FALSE, fig.align='center', fig.height=5, fig.width=6, fig.cap = cor_plot_fig_caption, message=FALSE, warning=FALSE}
cor_plot 
```

\newpage

```{r fac-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=6.5, fig.cap = fac_plot_fig_caption, message=FALSE, warning=FALSE}
f_ac_plot 
```

\newpage 

```{r mac-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=6.5, fig.cap = mac_plot_fig_caption, message=FALSE, warning=FALSE}
m_ac_plot 
```

\newpage

```{r icc-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=6.5, fig.cap = icc_plot_fig_caption, message=FALSE, warning=FALSE}
icc_plot 
```


\newpage
\begin{landscape}
```{r fpr-plot, echo=FALSE, fig.align='center', fig.height=5.5, fig.width=9, fig.cap = fpr_plot_fig_caption, message=FALSE, warning=FALSE}
fpr_plot
```

\newpage

```{r mpr-plot, echo=FALSE, fig.align='center', fig.height=5.5, fig.width=9, fig.cap = mpr_plot_fig_caption, message=FALSE, warning=FALSE}
mpr_plot
```
\end{landscape}

```{r ckmrsamp-plot, echo=FALSE, fig.align='center', fig.height=6.5, fig.width=6.5, fig.cap = ckmrsamp_plot_fig_caption, message=FALSE, warning=FALSE}
ckmr_sampling_plot
```

```{r sbys-plot1, echo=FALSE, fig.align='center', fig.height=7, fig.width=6.5, fig.cap = sbys_plot1_fig_caption, message=FALSE, warning=FALSE}
male_samps_suppl_plot1
```

```{r sbys-plot2, echo=FALSE, fig.align='center', fig.height=5, fig.width=6.5, fig.cap = sbys_plot2_fig_caption, message=FALSE, warning=FALSE}
male_samps_suppl_plot2
```


