---
title             : "Application of Close-kin mark–recapture (CKMR) to Protogynous Fishes"
subtitle          : "Supplementary Material"
author            : 
  - Claudia Friess, Eric C. Anderson, Lisa E. Ailloud, David S. Portnoy, and
  - Susan K. Lowerre-Barbieri

bibliography      : ["citations_abbrev.bib", "r-references.bib"]
csl               : "elsevier-harvard.csl"

floatsintext      : yes
linenumbers       : no
draft             : no
mask              : no

figurelist        : no
tablelist         : no
footnotelist      : no
always_allow_html : true

documentclass     : article
output:
  pdf_document:
    keep_tex: true
geometry          : margin=1in
linestretch       : 1.5
fontsize          : 12pt
header-includes   : 
  - \raggedright 
  - \usepackage{color}
  - \usepackage{tipa}
  - \usepackage{wasysym}
  - \usepackage{fancyhdr}
  - \usepackage{mathrsfs}
  - \thispagestyle{empty}
  - \usepackage{hanging}
  - \usepackage{setspace}
  - \usepackage{caption}
  - \usepackage{pdflscape}
  - \setlength{\baselineskip}{1.5\baselineskip}
  - \newcommand{\tprob}{P^{F\rightarrow M}}
  - \newcommand{\wt}[1]{\omega_#1}  % recruitment deviations
  - \newcommand{\rbar}{\bar{R}}  % mean recruitment
  - \newcommand{\fbarr}{{\bar{F}_\mathrm{rec}}}  % mean fishing mortality from rec fishery
  - \newcommand{\frecdev}[1]{fdev^\mathrm{rec}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\fbarc}{{\bar{F}_\mathrm{com}}}  % mean fishing mortality from commercial fishery
  - \newcommand{\fcomdev}[1]{fdev^\mathrm{com}_#1}  % annual deviations from mean mortality from rec fishery
  - \newcommand{\Mage}[1]{\dot{M}_#1}  % age-specific instantaneous natural mortality
  - \newcommand{\vulr}[1]{\dot{v}^\mathrm{rec}_#1}  % age-specific vulnerability to recreational fishery
  - \newcommand{\vulc}[1]{\dot{v}^\mathrm{com}_#1}  % age-specific vulnerability to commercial fishery
  - \newcommand{\dFec}[1]{\dot{Fec}_#1}    % age-specific relative fecundity
  - \newcommand{\Af}[1]{\dot{Mat}_#1}    % age-specific fraction of females mature
  - \newcommand{\Pfm}[1]{\dot{P}^{\mathrm{F}\rightarrow\mathrm{M}}_#1}    % age-specific prob of transitioning from female to male
  - \newcommand{\Imp}[1]{\dot{P}^\mathrm{M}_#1}    % equilibrium proportion of males amongst age a fish
  - \newcommand{\It}[1]{\mathring{I}_{#1}}   % recruitment index in year t
  - \newcommand{\pacom}[2]{\mathring{p}^\mathrm{com}_{#1,#2}}   % relative abundance of age a females in year t
  - \newcommand{\pacr}[3]{\mathring{p}^\mathrm{ckmr}_{#1,#2,#3}}   % relative abundance of age a males in year t
  - \newcommand{\Z}[2]{Z_{#1,#2}}   % total instantaneous mortality at time t for age a fish 
  - \newcommand{\AS}[2]{S_{#1,#2}}   % total annual survival
  - \newcommand{\lxf}[1]{s_{#1}}     % survivorship
  - \newcommand{\Nf}[2]{N^\mathrm{fem}_{{#1}, {#2}}}  % number of females of age a at time t
  - \newcommand{\Nm}[2]{N^\mathrm{mal}_{{#1}, {#2}}}  % number of males of age a at time t
  - \newcommand{\K}[4]{\mathring{K}^{\mathrm{#1},{{#4}}}_{{#2},{#3}}}
  - \newcommand{\KHS}[3]{\mathring{K}^{\mathrm{HSP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSs}[2]{\mathring{K}^{\mathrm{HSP},{{#2}}}_{{#1}}}
  - \newcommand{\KPOPT}[3]{\mathring{K}^{\mathrm{POP},{{#3}}}_{{#1},{#2}}}
  - \newcommand{\KHSD}[2]{\mathring{K}^{\mathrm{HSP}^{\neq _{a,b}}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with different mtDNA 
  - \newcommand{\KHSS}[2]{\mathring{K}^{\mathrm{HSP}^{=_a}}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOP}[3]{\mathring{K}^{\mathrm{POP}^{#3}}_{{#1},{#2}}}  
  - \newcommand{\KPOPF}[2]{\mathring{K}^\mathrm{POP_F}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA
  - \newcommand{\KPOPM}[2]{\mathring{K}^\mathrm{POP_M}_{{#1},{#2}}}   % number of half-sibling pairs (born c1 and c2) with same mtDNA 
  - \newcommand{\HSPD}{\mathrm{HSP}^{\neq}}
  - \newcommand{\HSPS}{\mathrm{HSP}^=}
  - \newcommand{\POPD}{\mathrm{POP}^{\neq}}
  - \newcommand{\POPS}{\mathrm{POP}^=}
  - \newcommand{\KNot}[2]{\mathring{K}^\mathrm{Not}_{{#1},{#2}}}     % number of pairs that are neither HSD or HSS
  - \newcommand{\KNote}[2]{\mathring{K}^{\mathrm{Not}=_a}_{{#1},{#2}}}
  - \newcommand{\KNotu}[2]{\mathring{K}^{\mathrm{Not}{\neq}_{a,b}}_{{#1},{#2}}}   
  - \newcommand{\Hmt}{H_\mathrm{mt}}  % mtDNA haplotype diversity
  - \newcommand{\Hmts}{H_\mathrm{mt}^{\mathrm{OM}}}  % simulated mtDNA haplotype diversity
  - \newcommand{\Hmte}{H_\mathrm{mt}^{\mathrm{EM}}}  % mtDNA haplotype diversity assumed in the EM
  - \newcommand{\Mat}{\mathrm{Mat}}
  - \newcommand{\Fec}{\mathrm{Fec}}
  - \DeclareMathAlphabet{\mathbbold}{U}{bbold}{m}{n}
  - \newcommand{\comment}[1]{\textcolor{red}{#1}}
  - \newcommand{\ERRO}{\mathrm{ERRO}}
  - \newcommand{\ETRO}{\mathrm{ETRO}}
  - \newcommand{\sH}{\mathscr{H}}
---

```{r setup, include = FALSE}
library(tidyverse)
library(stargazer)
library(kableExtra)
library(gridExtra)
library(nplyr)
library(patchwork)
library(ggh4x)
library(cowplot)
library(gridExtra)
library(grid)
library(scales)
library(here)

# caching for quick knitting.  
#    set cache = FALSE to not cache
#    set cache.extra to a new value to remove all caches and re-evaluate
knitr::opts_chunk$set(cache.extra = 4, cache = FALSE) 

source(here("code", "analysis", "results_suppl.R"), local = knitr::knit_global())


```


\renewcommand{\thefigure}{S\arabic{figure}} \setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}} \setcounter{table}{0}
\renewcommand{\theequation}{S\arabic{equation}} \setcounter{equation}{0}



## Female Breeding Group Switching in the Operating model

Each year, reproducing females spawned with 
their birth breeding group with probability $\zeta$, interpreted as their spawning 
group fidelity, set to 0.01. Females that switched breeding groups in a given breeding season 
(determined through Bernoulli draws with probability 1 - $\zeta$) switched from their 
birth breeding group $b_0$ to a different group $b_x$ with probability $\psi^{b_0,b_j}$. 
Probabilities $\psi$ were computed as the normalized inverse Euclidean distance 
values of a vector consisting of consistent intervals and having the length of 
the number of distinct breeding groups. This resulted in greater likelihoods of 
moving to a neighboring breeding group than a more distant one. 

## Assumption of Equifrequent Haplotypes in the Simulations

For the purposes of the simulations we present in this paper, we assume that all
mtDNA haplotypes are at the same frequency, (i.e., $f_a = 1/n_H$, for all $a \in \sH$, where
$n_H$ is the number of haplotypes in the set of all mtDNA haplotypes $\sH$). By doing so,
the simulations are greatly
simplified, as it is not necessary to track individual mtDNA haplotypes over generations
across all individuals. Rather, it is sufficient to record whether a sampled pair
shares the same mtDNA sequence or not. Furthermore, this allows us
to summarize the genetic variation in the mtDNA in terms of a single number, 
$\Hmt$, the probability that two mtDNA sequences randomly
sampled from the population are different in sequence.  

When mtDNA haplotypes are assumed equifrequent, we are able to categorize---without
loss of information---sampled pairs not by their specific
haplotypes, $a$ and $b$, but rather by whether
the pair shares the same haplotype or not, regardless of the specific sequence.
Thus in the CKMR pseudolikelihood to analyze the simulations we used joint probabilities
of a pair's type of relationship and whether or not the
members of the pair have the same mtDNA sequence. This joint probability is easily found
by summing the probabilities in (8) and (14).

Summing the first case ($a=b$) of (8) over all haplotypes $a\in\sH$ yields

\begin{equation}
\begin{aligned}
P(\mathrm{HSP}, \mathrm{mt}^=|c_i,c_j) &= P(\mathrm{HSP}_{F\rightarrow F}|c_i, c_j) \\
&\qquad + (1-\Hmt) \biggl[P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j) + P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j)\biggr]
\end{aligned}
\end{equation}
where $\mathrm{mt}^=$ denotes simply that the pair have the same mtDNA.  
Summing the
second case ($a\neq b$) in (8) over all pairs 
$(a,b),\ a\in\sH,\ b\in\sH,\ a\neq b$, gives

\begin{equation}
\begin{aligned}
P(\mathrm{HSP}, \mathrm{mt}^{\neq}|c_i,c_j) &= \Hmt\biggl[P(\mathrm{HSP}_{F\rightarrow M}|c_i, c_j) + P(\mathrm{HSP}_{M\rightarrow M}|c_i, c_j)\biggr].
\end{aligned}
\end{equation}
where $\mathrm{mt}^{\neq}$ denotes that the haplotypes carried by the pair's members
are of different sequences. 


Similarly, for POPs, summing (14) over different haplotypes provides the
simplification:

\begin{equation}
\begin{aligned}
P(\mathrm{POP}, \mathrm{mt}^= | c_i, c_j, g_i=F, s_i) &=  P(\mathrm{POP}_{F\rightarrow F}|c_i, c_j, s_i)\\
P(\mathrm{POP}, \mathrm{mt}^{\neq} | c_i, c_j, g_i=F, s_i) &= 0 \\
P(\mathrm{POP}, \mathrm{mt}^=|c_i, c_j, g_i=M, s_i) &=  P(\mathrm{POP}_{F\rightarrow M}|c_i, c_j, s_i) \\
& + (1-H_\mathrm{mt}) P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i) \\
P(\mathrm{POP}, \mathrm{mt}^{\neq}|c_i, c_j, g_i=M, s_i) &=  H_\mathrm{mt} P(\mathrm{POP}_{M\rightarrow M}|c_i, c_j, s_i) \\
\end{aligned}
\end{equation}

Both (S2) and (S3)) are easily used in a CKMR pseudolikelihood for
pairs categorized according to whether the mtDNA of the members matches or not.

## Additional Performance Metrics

In addition to median Percent Relative Error (PRE), we also calculated Normalized Root Mean 
Squared Error (NRMSE) for every scenario $sc$ as an additional accuracy metric:

\begin{equation}
\mathrm{NRMSE_{sc}} = \frac{\sqrt{\frac{1}{n_{sc}} \sum_{i=1}^{n_{sc}} (\mathrm{SV}_i - \mathrm{EV}_i)^2}}{\frac{1}{n_{sc}} \sum_{i=1}^{n_{sc}} \mathrm{SV}_i},
\end{equation}

where $n_sc$ is the number of simulations within a scenario. 

The full set of performance metrics for
sex-specific abundance and sex ratio are shown in Table S2. The pattern
of NRMSE was generally similar to median PRE in identifying the scenarios with the lowest
accuracy. However, due to the absolute rather than signed nature of the accuracy measure, error for the
base OM and EM scenarios appeared higher for NRMSE than PRE across CKMR sampling scenarios. Additionally,
the parameter estimation option with the lowest accuracy within scenario sometimes differed between
the two metrics (Figures S12 and S13).

## Investigating the Subsampling Effect

We subsampled each of the simulated populations ten times to create multiple data sets per population
for analysis. This was done for computational efficiency (it takes several hours to generate each pedigree),
but it also presents an opportunity to investigate the effect of random chance due to simulated pedigree
as opposed to random chance due to subsampling each pedigree. The subsampling effect was
explored by calculating the intraclass correlation coefficient (ICC) using one-way random effects
models of $\mathrm{PRE_c}$ with simulated population as the grouping factor. 
A separate model was fitted for each sex and each case
(combination of scenario and parameter estimation configuration) for a total of 48 models fitted. A low ICC
indicates a high degree of CKMR sampling stochasticity contribution to total variation and little effect
of population history. Conversely, a high ICC means the subsamples from the same population tend to cluster
together and differ from the subsamples of another population, with some populations consistenly producing
high PREs and others consistently producing lower ones.

We found that ICC was highly variable across cases, ranging from near zero to 0.78 (Figure S14). Without
model misspecification, CKMR sampling stochasticity was the dominant driver of variability (ICC < 0.2). 
ICC became elevated for some model misspecification, particularly when $\mathrm{RV}_{a,M}$ was simulated
to be steeper in the OM than $\mathrm{RV}_{a,F}$ and when $\mathrm{RV}_{a,M}$ was estimated for the
$\Hmte < \Hmts$ scenario (Figures S14, S15, and S16).

## Effect of changing the number of samples collected for CKMR and the timeframe over which they were collected

We ran a number of supplementary scenarios where the model was configured the same way as for the Base
scenario except for CKMR sample collection. In addition to the to the $10\sqrt{N}$ sample size
heuristic, the model was also run with 50% and 150% of that base sample size. We also explored whether 
the time period over which sampling is spread out affects the number of related pairs found and 
EM performance by changing the number of sampling years from three to ten. In the longer sampling scenario,
roughly the same overall CKMR sample size was collected as for the Base scenario, 
it was just spread out over more years.

The number of CKMR pairs found was higher when sampling was conducted
over a shorter time period, and this was most pronounced for
POPs sharing mitochondrial DNA and least pronounced for HSPs and POPs
with different mitochondrial DNA (Table S1, Figure S4). Additionally, the number
of pairs found was about nine times higher between the low (50% $10\sqrt{N}$) and high (150% $10\sqrt{N}$) CKMR
sample size runs and about four times higher between the low and
base (100% $10\sqrt{N}$) runs (Table S1, Figure S4).

There was no difference
in median bias between three and ten years of CKMR sampling, and precision
was only slightly lower for ten years of sampling (Table S2, Figure S17A). 
Precision increased by a relatively small amount (28%) when sample
size was increased from 100% to 150% of $10\sqrt{N}$ and more significantly
(${\approx}67\%$) when the sample size increased from 50% to 100% of $10\sqrt{N}$ (Figure S17A).

The number of sampling years continued having only a minor impact on estimation
when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and/or $\mathrm{RV}_{a,M}$ were free to be estimated. 
Generally, $\mathrm{IQR_{c}}$ increased when
sampling was spread out over ten years (an exception was male estimation performance when both $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ and $\mathrm{RV}_{a,M}$ were estimated). The largest increase (by ${\approx}11\%$) was observed for
females at 50% of $10\sqrt{N}$ sampling when $P_a^{\mathrm{F}\rightarrow \mathrm{M}}$ was estimated
(Table S2, Figure S17B&D). CKMR sample size continued to
impact precision more than the median of the abundance estimates. $\mathrm{IQR_{c}}$ increased from 29% to 41% when sample
size decreased from 150 to 100% $10\sqrt{N}$ and from 69% to 90% when sample size dropped from 100% to 50% of $10\sqrt{N}$
(Table S2).

## Effect of changing the number of males sampled for CKMR

The number of adult males sampled for CKMR in our simulation study was about two orders of
magnitude smaller than the number of females sampled. This might beg the question what would
happen if we were to increase the proportion of males in the sample. We conducted additional
runs on the base and skipped spawning OM scenarios to explore this question. In one set of runs,
we excluded age 1 females from the sex ratio calculation used to partition the CKMR sample by sex,
which roughly doubled the number of males sampled (Figure S18A) and
in another set, we sampled no males at all. The result was that the number of $\mathrm{POP}^{\neq}$
found roughly doubled when the number of males sampled doubled and decreased to zero when no males
were sampled, but the other kinship types were fairly unaffected (Figure S18B).
For estimation performance, we found that varying the number of males sampled had no effect 
(Figure S19), further emphasizing the minimal impact of
$\mathrm{POPs}$ on our estimator's accuracy and precision. 

## Supplementary Tables


```{r tableS1, echo=FALSE, warning=FALSE, results='asis'}
tS1_cap <- "Median number of adult males and females sampled and number of observed related pairs found by 
scenario as well as a number of supplemental runs investigating the effect of increasing
the number of years over which CKMR samples are collected (Yrs) and varying the number of 
samples collected (Nsamps; proportion of base sample size, see text for details). The 
number in parentheses is the interquartile range. sHmt = simulated mitochondrial haplotype diversity,
eHmt = mitochondrial haplotype diversity assumed in the EM. In the Base scenario, there was no skipped spawning
and male reproductive value (mRV) was equal to that of the females. In the Base EM, simulated haplotype diversity (sHmt)
was equal to estimated (eHmt). POD = parent-offspring pairs with different mitochondrial DNA,
POS = parent-offspring pairs with matching mitochondrial DNA, HSD = half-sibling
pairs with different mitochondrial DNA, HSS = half-sibling pairs with matching
mitochondrial DNA."

tab_tex <- knitr::kable(
  samps_table_data, 
  booktabs  = TRUE, 
  caption   = tS1_cap, 
  longtable = TRUE
) %>%
  add_header_above(c(" " = 3, "Number Sampled" = 2, "CKMR Pairs Found" = 5)) %>%
  kable_styling(
    latex_options          = c("hold_position", "repeat_header"),
    position               = "left", 
    font_size              = 10,
    repeat_header_method   = "replace"
  ) %>%
  kableExtra::landscape() %>%
  kableExtra::row_spec(
    0,
    extra_latex = "\\captionsetup{justification=centering, width=\\textwidth}"
  )

# Remove all addlinespace lines
tab_tex <- gsub("\\\\addlinespace\n", "", tab_tex)

cat(tab_tex)

```


```{r tableS2, echo=FALSE, warning=FALSE, results='asis'}
tS2_cap <- "Error metrics by scenario. PRE = Percent Relative Error. Median PREs were first calculated
across years within simulation to obtain a single PRE value per iteration. The PRE and IQR
(interquartile range) values reported in this table represent the median and IQR, respectively, of
those median PREs, by indicated scenario. NRMSE = Normalized Root Mean Square Error. 
(normalized by the mean of simulated values). N represents the number of unique data sets analyzed
for each scenario (400 minus the runs which didn't converge). In the Base scenario, male reproductive value (mRV) was
equal to that of females, there was no sperm limitation, the sampling model collected 
$10\\sqrt{N}$ (BaseN) samples over three years, and the EM fixed haplotype diversity (Hmt), mRV, and the sex 
transition function (trans) at the true values."
tab_tex <- knitr::kable(error_metrics_full, 
                  booktabs = TRUE, 
                  caption = tS2_cap, 
                  longtable = TRUE,
                  linesep = c("", "", "","", "\\addlinespace") # this doesn't work
                  ) %>%
  add_header_above(c(" " = 3, "Females" = 3, "Males" = 3, "Sex Ratio" = 3)) %>%
  kable_styling(latex_options = c("hold_position", "repeat_header"),
                position = "left", font_size = 10,
                repeat_header_method = "replace") %>%
  kableExtra::landscape() %>%
  kableExtra::row_spec(0, extra_latex = "\\captionsetup{justification=centering, width=\\textwidth}")

# Remove all addlinespace lines
tab_tex <- gsub("\\\\addlinespace\n", "", tab_tex)

cat(tab_tex)

```


```{r fig_caps, echo=FALSE}

age_sched_fig_caption <- "Age schedules for natural mortality (dashed line), female probability of reproducing (solid line), 
and sex transition probability (dotted line) used in operating model simulations. These were taken from the gag stock assessment
(SEDAR, 2022).
We allowed the terminal age to act as a plus group, 
rather than killing all indivduals that exceed the maximum age (the default behavior of fishSim) but set the natural mortality for this
plus group high to avoid building up a large pool of old individuals."

vulnF_plot_fig_caption <- "Fishing-related parameter values used in the simulations.  A) commercial (solid line) and recreational (dotted line) fishing fleet vulnerability at age and B) commercial (solid line) and recreational (dotted line) fishing mortality."

nsamps_plot_fig_caption <- "Number of adult fish CKMR samples collected by sex for the base OM configuration (no skipped spawning, male and female reproductive value are equal) for different sample sizes spread out over three (red) or ten (blue) years of sampling. The percentages refer to multiples of the base sample size, $n_\\mathrm{base} = 10\\sqrt{N}$, see text for details. The box represents the
interquartile range (IQR), the line is the median, the whiskers extend to 1.5 IQR and outliers fall beyond that."

supplckmrpairs_plot_fig_caption <- "Number of CKMR pairwise relationships observed for the base OM configuration (no skipped spawning, 
male and female reproductive value are equal) for different sample sizes spread out over three (red) or ten (blue) years of sampling. 
The percentages refer to multiples of the base sample size, $n_\\mathrm{base} = 10\\sqrt{N}$, see text for details. 
$\\mathrm{POP}^{\\neq}$ = parent-offspring pairs with different mitochondrial DNA, $\\mathrm{POP}^=$ = parent-offspring pairs with matching mitochondrial DNA, 
$\\mathrm{HSP}^{\\neq}$ = half-sibling pairs with different mitochondrial DNA, $\\mathrm{HSP}^=$ = half-sibling pairs with matching mitochondrial 
DNA. Boxplot elements are as described for Figure S3."


supplckmrpairs2_plot_fig_caption <- "Number of CKMR pairwise relationships for three years of CKMR sampling at the base sample size, 
$n_\\mathrm{base} = 10\\sqrt{N}$, by EM scenario. $\\mathrm{POP}^{\\neq}$ = parent-offspring pairs with different mitochondrial DNA, $\\mathrm{POP}^=$ = parent-offspring pairs with matching mitochondrial DNA, 
$\\mathrm{HSP}^{\\neq}$ = half-sibling pairs with different mitochondrial DNA, $\\mathrm{HSP}^=$ = half-sibling pairs with matching mitochondrial 
DNA. Boxplot elements are as described for Figure S3."

parnll_plot_fig_caption <- "Estimated parameters by EM scenario (relationship types used and mitochondrial haplotype diversity conditions) and whether or not the sex transition function standard deviation, $\\mathrm{STF}_{SD}$ of the sex transition and male the exponent of the reproductive value 
(RV) functions are assumed known (fixed).
trans = sex transition function, mRV = male reproductive value. Y-axis for mean recruitment and Male RV function exponent truncated to improve readability. Boxplot elements are as described for Figure S3."

par_plot_fig_caption <- "Estimated parameters by OM scenario (simulated male contribution to reproductive success and skipped spawning) and whether or not the sex transition function standard deviation and male reproductive value are assumed known (fixed). trans = sex transition function, mRV = male reproductive value. Y-axis for mean recruitment and Male RV function exponent truncated to improve readability. Boxplot elements are as described for Figure S3."

cor_plot_fig_caption <- "Correlation between key estimated parameter pairs (from the ADMB-estimated correlation matrix for
each individual simulation) when both the sex transition function standard deviation and male reproductive value (RV) are
estimated. Fct = function. Boxplot elements are as described for Figure S3."

fac_plot_fig_caption <- "Comparison of accuracy metrics for female abundance estimates, by OM
and EM scenario, and whether the sex transition function and male reproductive value were fixed or 
estimated in the model. PRE = Percent Relative Error, NRMSE = Normalized Root Mean Squared Error. RMSE was normalized by dividing by the mean of simulated values."

mac_plot_fig_caption <- "Comparison of accuracy metrics for male abundance estimates, by OM
and EM scenario, and whether the sex transition function and male reproductive value were fixed or 
estimated in the model. PRE = Percent Relative Error, NRMSE = Normalized Root Mean Squared Error. RMSE was normalized by dividing by the mean of simulated values."

fpr_plot_fig_caption <- "Percent Relative Error for female abundance estimates by simulated population, EM and OM
scenario, and whether the sex transition function and male reproductive value were fixed or estimated. Notice
the different scales between panels. Two high outliers omitted for scenario 'est Hmt < sim Hmt'. Boxplot elements are as described for Figure S3."

mpr_plot_fig_caption <- "Percent Relative Error for male abundance estimates by simulated population, EM and OM
scenario, and whether the sex transition function and male reproductive value were fixed or estimated. Notice
the different scales between panels. Two high outliers omitted for scenario 'est Hmt < sim Hmt'. Boxplot elements are as described for Figure S3."

sbys_plot1_fig_caption <- "Number of individuals sampled (Panel A) and number of kin pairs found (Panel B) for
the Base and skipped spawning OM scenarios when varying the proportion of males in the adult CKMR sample. When all
ages are included in calculating true sex ratio used to partition the sample by sex (base male sampling; red), 
fewer males are sampled than when age 1s (who are all female) are excluded from the sex ratio calculation (green). 
Boxplot elements are as described for Figure S3."

sbys_plot2_fig_caption <- "Percent relative error in female (Panel A) and male (Panel B) abundance estimation
for the Base and skipped spawning OM scenarios, by parameter estimation scenario, when varying the proportion of 
males in the adult CKMR sample. When all
ages are included in calculating true sex ratio used to partition the sample by sex (base male sampling; red), 
fewer males are sampled than when age 1s (who are all female) are excluded from the sex ratio calculation (green).
Boxplot elements are as described for Figure S3."

prebyyearM_plot_fig_caption <- "Percent relative error by year for males."

prebyyearF_plot_fig_caption <- "Percent relative error by year for females."

predist_plot_fig_caption <- "Distribution of median PREs by case."

icc_plot_fig_caption <- "ICC plot"

ckmrsamp_plot_fig_caption <- "Percent relative error in estimated male and female abundance time series 
summarized across years, by number of CKMR sampling years (red = 3, blue = 10), CKMR sample size, 
and whether or not sex transition 
($P^{\\mathrm{F}\\rightarrow \\mathrm{M}}$) and male reproductive value 
($\\mathrm{RV}_{M}$) functions are were assumed known (fixed). Results are shown for the base OM scenario and without
any model misspecification. Y-axis truncated to improve readability. Boxplot elements are as described for Figure S3."

```

## Supplementary Figures

```{r age-sched-plot, echo=FALSE, fig.align='center', fig.height=3, fig.cap = age_sched_fig_caption, fig.width=5, message=FALSE, warning=FALSE}
age_schedule_plot
```

```{r vulnF-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=5, fig.cap = vulnF_plot_fig_caption, message=FALSE, warning=FALSE}
F_vul_plot
```

\newpage
\begin{landscape}
```{r prebyyearM-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=9, fig.cap = prebyyearM_plot_fig_caption, message=FALSE, warning=FALSE}
pre_by_year_males 
```

\newpage

```{r prebyyearF-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=9, fig.cap = prebyyearF_plot_fig_caption, message=FALSE, warning=FALSE}
pre_by_year_females 
```

\newpage

```{r predist-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=9, fig.cap = predist_plot_fig_caption, message=FALSE, warning=FALSE}
grid.arrange(
  arrangeGrob(p1, p2, p3, p4, p5, p6, ncol = 3,
              left = textGrob("Count", rot = 90, vjust = 1),
              bottom = textGrob("Percent Error"))) 
```

\end{landscape}



\newpage
```{r nsamps-plot, echo=FALSE, fig.align='center', fig.height=4.5, fig.width=6.5, fig.cap = nsamps_plot_fig_caption, message=FALSE, warning=FALSE}
nsamps_plot
```

\newpage

```{r supplckmrpairs-plot, echo=FALSE, fig.align='center', fig.height=4.5, fig.width=6.5, fig.cap = supplckmrpairs_plot_fig_caption, message=FALSE, warning=FALSE}
suppl_ckmr_pairs_plot
```

\newpage

```{r supplckmrpairs2-plot, echo=FALSE, fig.align='center', fig.height=4.5, fig.width=6.5, fig.cap = supplckmrpairs2_plot_fig_caption, message=FALSE, warning=FALSE}
suppl_ckmr_pairs_plot2
```

\newpage
\begin{landscape}
```{r parnll-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=9, fig.cap = parnll_plot_fig_caption, message=FALSE, warning=FALSE}
para_res_plot_nll 
```

\newpage

```{r par-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=9, fig.cap = par_plot_fig_caption, message=FALSE, warning=FALSE}
para_res_plot 
```
\end{landscape}
\newpage

```{r cor-plot, echo=FALSE, fig.align='center', fig.height=5, fig.width=6, fig.cap = cor_plot_fig_caption, message=FALSE, warning=FALSE}
cor_plot 
```

\newpage

```{r fac-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=6.5, fig.cap = fac_plot_fig_caption, message=FALSE, warning=FALSE}
f_ac_plot 
```

\newpage 

```{r mac-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=6.5, fig.cap = mac_plot_fig_caption, message=FALSE, warning=FALSE}
m_ac_plot 
```

\newpage

```{r icc-plot, echo=FALSE, fig.align='center', fig.height=6, fig.width=6.5, fig.cap = icc_plot_fig_caption, message=FALSE, warning=FALSE}
icc_plot 
```


\newpage
\begin{landscape}
```{r fpr-plot, echo=FALSE, fig.align='center', fig.height=7.5, fig.width=10, fig.cap = fpr_plot_fig_caption, message=FALSE, warning=FALSE, out.width = "8.5in", out.height="6.375in"}
fpr_plot
```

\newpage

```{r mpr-plot, echo=FALSE, fig.align='center', fig.height=7.5, fig.width=10, fig.cap = mpr_plot_fig_caption, message=FALSE, warning=FALSE, out.width = "8.5in", out.height="6.375in"}
mpr_plot
```
\end{landscape}

```{r ckmrsamp-plot, echo=FALSE, fig.align='center', fig.height=7, fig.width=7, fig.cap = ckmrsamp_plot_fig_caption, message=FALSE, warning=FALSE}
ckmr_sampling_plot
```

```{r sbys-plot1, echo=FALSE, fig.align='center', fig.height=8, fig.width=7, fig.cap = sbys_plot1_fig_caption, message=FALSE, warning=FALSE}
male_samps_suppl_plot1
```

```{r sbys-plot2, echo=FALSE, fig.align='center', fig.height=5, fig.width=7, fig.cap = sbys_plot2_fig_caption, message=FALSE, warning=FALSE}
male_samps_suppl_plot2
```


